### 2021-05-04 paper/paper.tex
Switch to TEX

\documentclass[sigplan,review,screen]{acmart}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Type inference for the pi calculus with linear and shared channel types.

\textbf{TODO: recursion/replication} \textbf{TODO: bounded linear
types?} \textbf{TODO: type assertions}

Consider a \(\pi\)-calculus with usage annotations on channels. These
usage annotations are one of \(0\), \(1\), or \(\omega\). \(0\) must not
be used, \(1\) must be used exactly once, \(\omega\) can be used as many
times as desired. Sum and products contain usage annotations on their
leafs.

\hypertarget{types}{%
\subsection{Types}\label{types}}

Let \(\gamma\) be a context of metavariable kinds, where a kind is
either a \texttt{type} or a \texttt{usage}. Then \(\gamma \ni k\) is a
choice of a \(k\) inside \(\gamma\). \[
\begin{aligned}
S, T : \texttt{Type}~\gamma &= \texttt{mvar}~(\gamma \ni \texttt{type}) ~|~ \#_{i,o}t ~|~ \top ~|~ S + T ~|~ S \times T \\
i,o : \texttt{Usage}~\gamma &= \texttt{mvar}~(\gamma \ni \texttt{usage}) ~|~ 0\cdot ~|~ 1\cdot ~|~ \omega\cdot
\end{aligned}
\]

\hypertarget{type-system}{%
\subsection{Type System}\label{type-system}}

The type system is the usual standard type system based on context
splits that one defines for the pi calculus. TODO: insert type system
definition here:.

Usage annotations are \(0\cdot\), \(1\cdot\), \(\omega\cdot\). Context
splits are defined as follows:

\[
\begin{aligned}
\forall x.    && x           &= x + 0\cdot \\
\forall x.    && x           &= 0\cdot + x \\
\forall x ~y. && \omega\cdot &= y + z
\end{aligned}
\]

Only the usage annotations at the upper level of a channel type are
split: the usage annotations of any channel type that is send as data
are preserved on all three sides of the split.

\hypertarget{inference}{%
\section{Inference}\label{inference}}

Co-contextual type inference: traverses the syntax bottom up, collecting
constraints on the typing context.

Constraints can be of the following form:

\begin{itemize}
\item
  \(S = T\): types \(S\) and \(T\) must be unifiable.
\item
  \(S = T + R\): adding up the usage annotations in \(T\) and \(R\) must
  result in \(S\).
\end{itemize}

Constraints of the form \(S = T\) can always be eagerly unified. However
constraints of the form \(S = T + R\) might contain variables:
\(\texttt{mvar}~x = \texttt{mvar}~y + \texttt{mvar}~z\). In an open
process, these variables cannot safely be instantiated to \(0\). As
such, no substitution is possible, yet we need to remember the
constraint. (It might also be that some constraints have more than one
possible solution.) At the same time, we need to show that the
constraints that we postulate are enough to solve the problem.

\paragraph{Constraint Satisfaction}

The \([\![\_]\!]\) function interprets constraints into their type
counterparts \(S \equiv T\) and \(S = T + R\). Substituting into a
constraint substitutes pointwise.

\paragraph{Inference}

\[
\texttt{infer} : \forall (p : \texttt{Proc}~n). ~ \exists \gamma. ~ \texttt{Ctx}~n~\gamma \times \texttt{List}~(\texttt{Constr}~\gamma)
\]

\paragraph{Inference soundness.}

Every substitution \(\sigma\) that makes the constraints hold will make
the process typable under the substituted context. \[
\begin{aligned}
\texttt{infer-sound} &: \forall (p : \texttt{Proc}~n). ~ \texttt{infer}~P \equiv \gamma , \Gamma , cs \\
&\to \forall (\sigma : \texttt{Subst}~\gamma~\delta). ~ [\![ \sigma \triangleleft cs ]\!] \\
&\to (\sigma \triangleleft \Gamma) \vdash P
\end{aligned}
\]

\paragraph{Inference completeness.}

For every context \(\Delta\) that makes the process typable there exists
a most general substitution \(\sigma\) that will solve the constraints
and which substitutes in \(\Gamma\) to something more general than
\(\Delta\). \[
\begin{aligned}
\texttt{infer-complete} &: \forall (p : \texttt{Proc}~n). ~ \texttt{infer}~P \equiv \gamma , \Gamma , cs \\
&\to \forall (\Delta : \texttt{Ctx}~n~\gamma). ~ \Delta \vdash P \\
&\to \exists \delta. ~ \exists (\sigma : \texttt{Subst}~\gamma~\delta). \\
&\to [\![ \sigma \triangleleft cs ]\!] \times \Delta \subseteq (\sigma \triangleleft \Gamma)
\end{aligned}
\] where \[
\Delta \subseteq \Gamma \triangleq \exists \delta. ~ \exists (\sigma : \texttt{Subst} \_ \delta). ~ \Delta \equiv (\sigma \triangleleft \Gamma)
\]

\hypertarget{constraint-resolution}{%
\section{Constraint Resolution}\label{constraint-resolution}}

\paragraph{Constraint Resolution.}

Given a set of accumulated substitutions and some constraints, we will
return (hopefully) some further substitutions, and a set of leftover
constraints to which the substitutions have been applied. These leftover
constraints can either not be simplified or are unsatisfiable. \[
\begin{aligned}
\texttt{solve} &: \texttt{Subst}~\gamma~\delta \times \texttt{List}~(\texttt{Constr}~\delta) \\
&\to \texttt{Subst}~\gamma~\phi \times \texttt{List}~(\texttt{Constr}~\phi)
\end{aligned}
\]

\paragraph{Soundness.}

The simplified constraints are enough to entail the original
constraints. \[
\begin{aligned}
\texttt{solve-sound}
&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
&\to \forall (\sigma_f : \texttt{Subst}~\gamma~\delta) \\
& \to [\![ \sigma_f \triangleleft cs_2 ]\!] \to [\![ \sigma_f \cdot \sigma_2 \triangleleft cs_1 ]\!]
\end{aligned}
\]

\paragraph{Completeness.}

Any substitution \(\sigma_f\) that makes the original constraints
\(cs_1\) hold will be a specialisation \(\sigma_f\) of our returned
substitution \(\sigma_2\). The specialisation \(\sigma_g\) will make the
returned constraints \(cs_2\) hold. \[
\begin{aligned}
\texttt{solve-sound}
&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
&\to \forall (\sigma_f : \texttt{Subst}~\gamma~\delta). ~ [\![ \sigma_f \triangleleft cs_1 ]\!] \\
&\to \exists (\sigma_g : \sigma_f = \sigma_g \cdot \sigma_2) \times [\![ \sigma_g \triangleleft cs_2 ]\!]
\end{aligned}
\]

\paragraph{Progress.}

Nothing prevents us from returning the original constraints as output.
We therefore promise that every constraint we return is either currently
unsolvable because we have not enough information, or is unsatisfiable
altogether. \[
\begin{aligned}
\texttt{solve-sound}
&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
&\to \forall c \in cs_2. ~ \texttt{IsDeferred}~c \uplus \texttt{IsUnsat}~c
\end{aligned}
\] where \[
\texttt{IsUnsat}~c = \forall \sigma. ~ \neg [\![ \sigma \triangleleft c ]\!]
\]

\paragraph{Instantiation.}

\[
\forall c. ~ \texttt{IsDeferred}~c \to \exists \sigma. [\![ \sigma \triangleleft c ]\!]
\]

\end{document}

### 2021-05-04 paper/paper.tex
Add abstract

@@
-\documentclass[sigplan,review,screen]{acmart}
+\documentclass[sigplan,screen,review]{acmart}
+
+\usepackage{todonotes}
+
+\newcommand{\picalc}{$\pi$-calculus}
+\newcommand{\Picalc}{$\pi$-Calculus}
+
+\title{Extended Abstract}
+\subtitle{Co-Contextual Type Inference for the Linear \Picalc{} in Agda}
+
+\author{Uma Zalakain}
+\affiliation{University of Glasgow}
+\email{u.zalakain.1@research.gla.ac.uk}
+
+\author{Ornela Dardha}
+\affiliation{University of Glasgow}
+\email{ornela.dardha@glasgow.ac.uk}
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\begin{abstract}
+  The linear \picalc{} is an important substructural type system for the \picalc{}:
+  it is a common target for the encoding of session-typed languages \todo{cite},
+  and it in direct correspondence with classical linear logic \todo{cite}.
+
+  Type inference for the linear \picalc{} comes with its own challenges:
+  for one, processes themselves are untyped, and thus typing constraints only apply to the typing contexts;
+  moreover, typing constraints on usage annotations cannot always be eagerly solved in open processes.
+
+  We approach the first challenge using co-contextual type inference, where processes are traversed bottom-up and constraints are collected (and when possible eagerly solved) and merged on context splits.
+  We approach the latter challenge by allowing the resolution of some of the constraints to be deferred, and adapting the soundness and completeness theorems to take this into account.
+  We mechanise all our work in Agda.
+\end{abstract}
+
+\maketitle
+
+\section{Introduction}\label{introduction}
+
+\cite{ErdwegBKKM15}
 
@@
 
+\bibliographystyle{abbrvnat}
+\bibliography{paper}
 \end{document}

### 2021-05-04 paper/paper.tex
Start working on intro

--- 2021-05-04 paper/paper.tex
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\begin{abstract}
+  The linear \picalc{} is a well-known substructural type system for the \picalc{}:
+  it is a common target for the encoding of session-typed languages \todo{cite},
+  and it in direct correspondence with classical linear logic \todo{cite}.
 
-Type inference for the pi calculus with linear and shared channel types.
+  Type inference for the linear \picalc{} comes with its own challenges:
+  for one, processes themselves are untyped, and thus typing constraints only apply to the typing contexts;
+  moreover, typing constraints on usage annotations \emph{cannot always be eagerly solved in open processes}.
+
+  We approach the first challenge using \emph{co-contextual type inference} \cite{ErdwegBKKM15}, where processes are traversed bottom-up and constraints are collected (and when possible eagerly solved) and merged on context splits.
+  We approach the latter challenge by allowing the resolution of some of the constraints to be deferred.
+  We state soundness and completeness theorems that take this deferral of constraint resolution into account.
+  We state a progress (?) theorem that shows that only those constraints that need to be deferred are deferred.
+  \todo{be more precise}
+  We are working on mechanising our work in Agda.
+\end{abstract}
+
+\maketitle
+
+\section{Introduction}\label{introduction}
+
+We define an untyped but well scoped \picalc{} using type-level de Bruijn \todo{cite} indices \autoref{syntax}.
+We define a small expression language within the process syntax to handle composite sum and product types.
+On top of the well-scoped syntax we provide the standard linear type-system using context-splits \autoref{type-system}.
+The type system is indexed by a typing context that admits unsolved metavariables.
+These unsolved metavariables are however well kinded: either types or usage annotations.
+
+\todo{insert example showing why we need deferral}
+
+We split typing inference into constraint collection and constraint solving \autoref{inference}.
+We start by defining constraints on types and usage annotations \autoref{constraints}
+We collect constraints bottom-up, creating extra \emph{merging} constraints on context splits.
+This allows type-inference to be easily parallelisable.
+We postulate soundness and completeness theorems that allow for constraint solving to be deferred \autoref{inference-theorems}.
 
@@
 
-\hypertarget{types}{%
-\subsection{Types}\label{types}}
+\section{Type System}\label{type-system}
+
+\subsection{Types}\label{types}
 
@@
 \]
-
-\hypertarget{type-system}{%
-\subsection{Type System}\label{type-system}}
 

### 2021-05-04 paper/paper.tex
Add type system

@@
 
+% Typing rules
+\usepackage{mathpartir}
+\mprset {sep=1em} % Horizontal space between premises
+
 \newcommand{\picalc}{$\pi$-calculus}
 \newcommand{\Picalc}{$\pi$-Calculus}
+\newcommand{\rulename}[1]{{\tiny \textsc{(#1)}}}
+\newcommand{\constr}[1]{\textcolor{olive}{\mathtt{#1}}}
+\newcommand{\type}[1]{\textcolor{blue}{\mathtt{#1}}}
+
+% Syntax types
+\newcommand{\Fin}[1]{\type{Fin}~#1}
+\newcommand{\sExpr}[1]{\type{Expr}~#1}
+\newcommand{\sProc}[1]{\type{Proc}~#1}
+\newcommand{\tvar}[2]{#1 ~\type{\ni_t}~ #2}
+\newcommand{\tkind}[2]{#1 ~\type{\vdash_t}~ #2}
+\newcommand{\ttype}[1]{\type{Type}~#1}
+\newcommand{\tusage}[1]{\type{Usage}~#1}
+\newcommand{\tCtx}[2]{\type{Ctx}_{#1}~#2}
+\newcommand{\tSplit}[3]{#1~\type{=}~#2~\type{\uplus}~#3}
+\newcommand{\tun}[1]{\type{un}~#1}
+\newcommand{\tVar}[3]{#1 ~ \type{\ni} ~ #2 ~ \type{:} ~ #3}
+\newcommand{\tExpr}[3]{#1 ~ \type{\vdash} ~ #2 ~ \type{:} ~ #3}
+\newcommand{\tProc}[2]{#1 ~ \type{\vdash} ~ #2}
+
+% Syntax constructors
+\newcommand{\sunit}{\constr{unit}}
+\newcommand{\svar}{\constr{var}}
+\newcommand{\sfst}{\constr{fst}}
+\newcommand{\ssnd}{\constr{snd}}
+\newcommand{\sinl}{\constr{inl}}
+\newcommand{\sinr}{\constr{inr}}
+\newcommand{\spair}{\constr{pair}}
+\newcommand{\send}{\constr{end}}
+\newcommand{\snew}{\constr{new}}
+\newcommand{\scomp}{\constr{comp}}
+\newcommand{\srecv}{\constr{recv}}
+\newcommand{\ssend}{\constr{send}}
+\newcommand{\scase}{\constr{case}}
+
+% Kind constructors
+\newcommand{\ktype}{\constr{ty}}
+\newcommand{\kusage}{\constr{us}}
+
+% Type constructors
+\newcommand{\tmvar}{\constr{mvar}}
+\newcommand{\tchan}{\constr{chan}}
+\newcommand{\tunit}{\constr{unit}}
+\newcommand{\tsum}{\constr{sum}}
+\newcommand{\tprod}{\constr{prod}}
+\newcommand{\tzero}{\constr{0\cdot}}
+\newcommand{\tone}{\constr{1\cdot}}
+\newcommand{\tomega}{\constr{\omega\cdot}}
 
@@
 
-\textbf{TODO: recursion/replication} \textbf{TODO: bounded linear
-types?} \textbf{TODO: type assertions}
+\section{Syntax}\label{syntax}
 
-Consider a \(\pi\)-calculus with usage annotations on channels. These
-usage annotations are one of \(0\), \(1\), or \(\omega\). \(0\) must not
-be used, \(1\) must be used exactly once, \(\omega\) can be used as many
-times as desired. Sum and products contain usage annotations on their
-leafs.
+Expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$, variable references $i_n$ are of type $\Fin{n}$.
+\[
+\begin{aligned}[c]
+  e_n ~ f_n  :=
+  &~ \sunit \\
+  |&~ \svar~i_n \\
+  |&~ \spair~e_n~f_n \\
+  |&~ \sfst~e_n ~|~  \ssnd~e_n \\
+  |&~ \sinl~e_n ~|~  \sinr~e_n
+\end{aligned}
+\begin{aligned}[c]
+  p_n ~ q_n  :=
+  &~ \send ~|~  \snew~p_{1+n} \\
+  |&~ \srecv~e_n~p_{1+n} \\
+  |&~ \ssend~e_n~f_n~p_n \\
+  |&~ \scomp~p_n~q_n \\
+  |&~ \scase~e_n~p_{1+n}~q_{1+n}
+\end{aligned}
+\]
+\todo{add recursion}
+\todo{add type assertions}
 
@@
 
-\subsection{Types}\label{types}
+Both types and usage annotations allow for \emph{holes} or \emph{metavariables}.
+We use the same set of metavariables for both types and usage annotations -- this makes their handling more uniform.
+We use a context of kinds \(\gamma\) to keep track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$: a choice of kind \(k\) in a kinding context \(\gamma\) is denoted as $\tvar{\gamma}{k}$.
 
-Let \(\gamma\) be a context of metavariable kinds, where a kind is
-either a \texttt{type} or a \texttt{usage}. Then \(\gamma \ni k\) is a
-choice of a \(k\) inside \(\gamma\). \[
-\begin{aligned}
-S, T : \texttt{Type}~\gamma &= \texttt{mvar}~(\gamma \ni \texttt{type}) ~|~ \#_{i,o}t ~|~ \top ~|~ S + T ~|~ S \times T \\
-i,o : \texttt{Usage}~\gamma &= \texttt{mvar}~(\gamma \ni \texttt{usage}) ~|~ 0\cdot ~|~ 1\cdot ~|~ \omega\cdot
-\end{aligned}
-\]
+We define well kinded types and usage annotations simultaneously so that we are able to provide a uniform unification and context merge algorithm later on (note that we define $\tsum$ and $\tprod$ and $\tzero$, $\tone$ and $\tomega$ in one go to save space):
+\begin{mathpar}
+  \inferrule {m : \tvar{\gamma}{k}} {\tmvar~m : \tkind{\gamma}{k}}
 
-The type system is the usual standard type system based on context
-splits that one defines for the pi calculus. TODO: insert type system
-definition here:.
+  \inferrule {
+    i : \tkind{\gamma}{\kusage} \\
+    o : \tkind{\gamma}{\kusage} \\
+    t : \tkind{\gamma}{\ktype}}
+  {\tchan~i~o~t : \tkind{\gamma}{\ktype}}
 
-Usage annotations are \(0\cdot\), \(1\cdot\), \(\omega\cdot\). Context
-splits are defined as follows:
+  \inferrule { }
+  {\tunit : \tkind{\gamma}{\ktype}}
 
-\[
-\begin{aligned}
-\forall x.    && x           &= x + 0\cdot \\
-\forall x.    && x           &= 0\cdot + x \\
-\forall x ~y. && \omega\cdot &= y + z
-\end{aligned}
-\]
+  \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
+  {\tsum~s~t ~|~ \tprod~s~t : \tkind{\gamma}{\ktype}}
 
-Only the usage annotations at the upper level of a channel type are
-split: the usage annotations of any channel type that is send as data
-are preserved on all three sides of the split.
+  \inferrule { } {\tzero ~|~ \tone ~|~ \tomega : \tkind{\gamma}{\kusage}}
+\end{mathpar}
+We will henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
 
-\hypertarget{inference}{%
-\section{Inference}\label{inference}}
+A context $\Gamma$ of type $\tCtx{n}{\gamma}$ is a list of $\ttype{\gamma}$ of size $n$.
+We define context splits $\tSplit{\Gamma}{\Delta}{\Theta}$ pointwise on types.
+Splits on types are defined pointwise recursively on usage annotations -- note however that only the usage annotations on the top level of a channel (that is, not within its payload) are split.
+Splits on usage annotations are defined as follows -- note that they are not necessarily unique:
+\begin{mathpar}
+  \inferrule { } {\tSplit{x}{x}{\tzero}}
+
+  \inferrule { } {\tSplit{x}{\tzero}{x}}
+
+  \inferrule { } {\tSplit{\tomega}{x}{y}}
+\end{mathpar}
+
+A usage annotation $x$, type $t$ or context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $x$ and $t$) \todo{cite padovani}.
+\todo{typing judgment for variables}
+
+We can finally define our typing judgment for expressions and processes. \todo{formation rules double lined}
+
+\begin{mathpar}
+  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ e : \sExpr{n} \\ t : \ttype{\gamma}}
+             { \tExpr{\Gamma}{e}{t} }
+
+  \inferrule { \tVar{\Gamma}{i}{t} } { \tExpr{\Gamma}{\svar~i}{t} } \rulename{var}
+
+  \inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} } \rulename{unit}
+
+  \inferrule { \tExpr{\Gamma}{e}{\tprod~t~s} \\ \tun{s}} { \tExpr{\Gamma}{\sfst~e}{t}} \rulename{fst}
+
+  \inferrule { \tExpr{\Gamma}{e}{t} } { \tExpr{\Gamma}{\sinl~e}{\tsum~t~s}} \rulename{inl}
+
+  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{s} \\ \tExpr{\Theta}{f}{t} }
+             { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
+             \rulename{pair}
+
+  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ p : \sProc{n}}
+             { \tProc{\Gamma}{p} }
+
+  \inferrule { \tun{\Gamma} } { \tProc{\Gamma}{\send} } \rulename{end}
+
+  \inferrule { t : \ttype{\gamma} \\ \tProc{\Gamma , t}{p} } { \tProc{\Gamma}{\snew~p} }
+
+  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tchan~\tone~\tzero~t} \\ \tProc{\Theta,t}{p} }
+             { \tProc{\Gamma}{\srecv~e~p} }
+
+  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tsum~s~t} \\ \tProc{\Theta,s}{p} \\ \tProc{\Theta,t}{q} }
+             { \tProc{\Gamma}{\scase~e~p~q} }
+\end{mathpar}
+
+\section{Inference}\label{inference}
 

### 2021-05-10 paper/paper.tex
Add main references

@@
   The linear \picalc{} is a well-known substructural type system for the \picalc{}:
-  it is a common target for the encoding of session-typed languages \todo{cite},
+  it is a common target for the encoding of session-typed languages \cite{DardhaGS17},
   and it in direct correspondence with classical linear logic \todo{cite}.
@@
 
-We define an untyped but well scoped \picalc{} using type-level de Bruijn \todo{cite} indices \autoref{syntax}.
+We define an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices \autoref{syntax}.
 We define a small expression language within the process syntax to handle composite sum and product types.
@@
 
-A usage annotation $x$, type $t$ or context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $x$ and $t$) \todo{cite padovani}.
+A usage annotation $x$, type $t$ or context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $x$ and $t$) \cite{Padovani15}.
 \todo{typing judgment for variables}
@@
 
+\cite{McBride03}
+
 \paragraph{Constraint Resolution.}

### 2021-05-10 paper/paper.tex
Add variable typing rules
Write intro to inference

@@
 \newcommand{\Fin}[1]{\type{Fin}~#1}
+\newcommand{\fzero}{\constr{zero}}
+\newcommand{\fsuc}{\constr{suc}}
 \newcommand{\sExpr}[1]{\type{Expr}~#1}
@@
 
-% Type constructors
+% Typing judgment constructors
 \newcommand{\tmvar}{\constr{mvar}}
@@
 \newcommand{\tomega}{\constr{\omega\cdot}}
+
+% Constraint constructors
+\newcommand{\eqconstr}[2]{#1 ~ \constr{\stackrel{c}{=}} ~ #2}
+\newcommand{\sumconstr}[3]{#1 ~ \constr{\stackrel{c}{=}} ~ #2 ~ \constr{+} ~ #3}
 
@@
 A usage annotation $x$, type $t$ or context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $x$ and $t$) \cite{Padovani15}.
-\todo{typing judgment for variables}
 
-We can finally define our typing judgment for expressions and processes. \todo{formation rules double lined}
+We can finally define our typing judgment for variables, expressions and processes -- some are omitted for brevity.
 
 \begin{mathpar}
-  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ e : \sExpr{n} \\ t : \ttype{\gamma}}
-             { \tExpr{\Gamma}{e}{t} }
+%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ i : \Fin{n} \\ t : \ttype{\gamma} }
+%             { \tVar{\Gamma}{i}{t} }
+
+  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ \tun{\Gamma} \\ t : \ttype{\gamma} }
+             { \tVar{\Gamma,t}{\fzero}{t} }
+             \rulename{zero}
+
+  \inferrule { \tVar{\Gamma}{i}{t} \\ s : \ttype{\gamma} \\ \tun{s} }
+             { \tVar{\Gamma,s}{\fsuc~i}{t} }
+             \rulename{suc}
+
+%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ e : \sExpr{n} \\ t : \ttype{\gamma}}
+%             { \tExpr{\Gamma}{e}{t} }
 
@@
 
-  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ p : \sProc{n}}
-             { \tProc{\Gamma}{p} }
+%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ p : \sProc{n}}
+%             { \tProc{\Gamma}{p} }
 
@@
   \inferrule { t : \ttype{\gamma} \\ \tProc{\Gamma , t}{p} } { \tProc{\Gamma}{\snew~p} }
+             \rulename{new}
 
@@
              { \tProc{\Gamma}{\srecv~e~p} }
+             \rulename{recv}
 
@@
              { \tProc{\Gamma}{\scase~e~p~q} }
+             \rulename{case}
 \end{mathpar}
@@
 
-Co-contextual type inference: traverses the syntax bottom up, collecting
-constraints on the typing context.
+Co-contextual type inference traverses processes bottom up collecting constraints.
+Limiting the information flow to being bottom up has the advantage of making inference subproblems independent and thus easily parallelisable.
+For the shared \picalc{}, inference constraints amount to unification problems that can always be eagerly solved.
+The linear \picalc{} however requires that some constraints are deferred to a later stage, when more information is available.
 
-Constraints can be of the following form:
+Let us suppose we can safely instantiate to $\tzero$ every usage annotation that is unrestricted.
+Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we send the unit over $a$, then send $a$ over $x$ and terminate.
+By the time we reach $\send$ both $x$ and $a$ must be unrestricted, and thus channel $x$ must be of some type $\tchan~\tzero~\tone~t$.
+Since we send a unit over $a$, $a$ must be of some type $s$ such that $\tSplit{s}{\tchan~\tzero~\tone~\tunit}{t}$.
+From this we can infer that both $s$ and $t$ must be a channel type and thus $\tSplit{s_i}{\tzero}{t_i}$ and $\tSplit{s_o}{\tone}{t_o}$ for some $s_i$, $t_i$, $s_o$ and $t_o$, but we cannot infer their usage annotations: all we can deduce is their lower bounds $\tzero$ and $\tone$.
+We must defer the resolution of these constraints until more information becomes available about the context of this open process.
 
-\begin{itemize}
-\item
-  \(S = T\): types \(S\) and \(T\) must be unifiable.
-\item
-  \(S = T + R\): adding up the usage annotations in \(T\) and \(R\) must
-  result in \(S\).
-\end{itemize}
+We present two kinds of constraints, showcase how to encode inference problems for some of the typing rules in \autoref{type-system}, and state the lemmas that will allow us to show that our approach to inference is both sound and complete.
 
-Constraints of the form \(S = T\) can always be eagerly unified. However
-constraints of the form \(S = T + R\) might contain variables:
-\(\texttt{mvar}~x = \texttt{mvar}~y + \texttt{mvar}~z\). In an open
-process, these variables cannot safely be instantiated to \(0\). As
-such, no substitution is possible, yet we need to remember the
-constraint. (It might also be that some constraints have more than one
-possible solution.) At the same time, we need to show that the
-constraints that we postulate are enough to solve the problem.
+Constraints take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where the usage annotations in $T$ and $R$ are meant to be added up to $S$.
+Both these constraints are defined on usage annotations, types, and typing contexts.
+Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter \emph{sometimes} cannot, and must be deferred.
 

### 2021-05-10 paper/paper.tex
Add color to functions, note notation

@@
 \newcommand{\constr}[1]{\textcolor{olive}{\mathtt{#1}}}
+\newcommand{\func}[1]{\textcolor{gray}{\mathtt{#1}}}
 \newcommand{\type}[1]{\textcolor{blue}{\mathtt{#1}}}
@@
 \newcommand{\tSplit}[3]{#1~\type{=}~#2~\type{\uplus}~#3}
+\newcommand{\tEq}[2]{#1~\type{\equiv}~#2}
 \newcommand{\tun}[1]{\type{un}~#1}
@@
 We postulate soundness and completeness theorems that allow for constraint solving to be deferred \autoref{inference-theorems}.
+
+(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and this is not a poem.)
 
@@
 
+\paragraph{Substitution}
+
 \paragraph{Constraint Satisfaction}
 
-The \([\![\_]\!]\) function interprets constraints into their type
-counterparts \(S \equiv T\) and \(S = T + R\). Substituting into a
-constraint substitutes pointwise.
+We use the $\func{[\![\_]\!]}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.
+
+\todo{substitution into constraints}
 

### 2021-05-10 paper/paper.tex
A little on inference

--- 2021-05-10 paper/paper.tex
@@
 \newcommand{\tProc}[2]{#1 ~ \type{\vdash} ~ #2}
+\newcommand{\tConstr}[1]{\type{Constr} ~ #1}
+\newcommand{\tConstrs}[1]{\type{[Constr} ~ #1 ~ \type{]}}
 
@@
 \newcommand{\tomega}{\constr{\omega\cdot}}
+
+\newcommand{\subst}[2]{#1 ~ \func{\triangleleft} ~ #2}
 
@@
 We postulate soundness and completeness theorems that allow for constraint solving to be deferred \autoref{inference-theorems}.
+
+(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and I am sorry this is not a wee poem.)
 
@@
 
-Constraints take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where the usage annotations in $T$ and $R$ are meant to be added up to $S$.
-Both these constraints are defined on usage annotations, types, and typing contexts.
+\paragraph{Constraints}
+
+Constraints are of type $\tConstr{\gamma}$ and are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, both usage annotations and types.
+They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
 Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter \emph{sometimes} cannot, and must be deferred.
+We use $\tConstrs{\gamma}$ to refer to lists of constraints of type $\tConstr{\gamma}$.
 
-\paragraph{Constraint Satisfaction}
+\paragraph{Substitution}
 
-The \([\![\_]\!]\) function interprets constraints into their type
-counterparts \(S \equiv T\) and \(S = T + R\). Substituting into a
-constraint substitutes pointwise.
+We encode substitution as a function $\subst{\_}{\_}$ of type $(\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}) \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ that acts on usage annotations and types and typing contexts, and that freely alters the metavariable context.
+We define substitutions on constraints pointwise on their arguments.
+
+\paragraph{Constraint satisfaction}
+
+We use a $\func{[\![\_]\!]}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.
 
@@
 
-\[
-\texttt{infer} : \forall (p : \texttt{Proc}~n). ~ \exists \gamma. ~ \texttt{Ctx}~n~\gamma \times \texttt{List}~(\texttt{Constr}~\gamma)
-\]
+We encode type inference as a function taking processes with $n$ free variables and returning a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$:
+$$
+\func{infer} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma}
+$$
+
+\todo{insert examples}
+
+We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression.
+
+\todo{insert examples}
 

### 2021-05-10 paper/paper.tex
Start working on some inference examples

@@
 % Constraint constructors
-\newcommand{\eqconstr}[2]{#1 ~ \constr{\stackrel{c}{=}} ~ #2}
-\newcommand{\sumconstr}[3]{#1 ~ \constr{\stackrel{c}{=}} ~ #2 ~ \constr{+} ~ #3}
+\newcommand{\eqconstr}[2]{\constr{[} ~ #1 ~ \constr{\stackrel{c}{=}} ~ #2 ~ \constr{]}}
+\newcommand{\sumconstr}[3]{\constr{[} ~ #1 ~ \constr{\stackrel{c}{=}} ~ #2 ~ \constr{+} ~ #3 ~ \constr{]}}
 
@@
 $$
-\func{infer} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma}
+\func{inferProc} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma}
 $$
+We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression.
+$$
+\func{inferExpr} : \sExpr{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \ttype{\gamma} \times \tConstrs{\gamma}
+$$
+Let us provide a couple of examples illustrating how this function is defined:
+\begin{itemize}
+\item
+  For the expression $\sfst~e$ we first recursively infer the metavariable context $\gamma$, the typing context $\Gamma$, the type $\tau$ and the constraints $Cs$ for $e$.
+  We then add two fresh metavariables $?t$ and $?s$ of kind $\ktype$ to $\gamma$, add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $cs$, and return $?t$ as the type of $e$.
 
-\todo{insert examples}
-
-We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression.
-
-\todo{insert examples}
+\item
+  For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \gamma_p, (\Gamma_p, s) , Cs_p$, $\func{inferProc}~q = \gamma_q, (\Gamma_q, t) , Cs_q$, and $\func{inferExpr}~e = \gamma_e, \Gamma_e, \tau, Cs_e$.
+  We add the metavariables in $\gamma_p$ to the ones in $\gamma_q$ and a new metavariable $?r$ of kind $\ktype$, and refer to the result as $\delta$.
+  We add the constraints in $Cs_p$  to the ones in $Cs_q$ and create new constraints $\eqconstr{\Gamma_p}{\Gamma_q}$ 
+\end{itemize}
 

### 2021-05-11 paper/paper.tex
Finish example inference

@@
 \item
-  For the expression $\sfst~e$ we first recursively infer the metavariable context $\gamma$, the typing context $\Gamma$, the type $\tau$ and the constraints $Cs$ for $e$.
-  We then add two fresh metavariables $?t$ and $?s$ of kind $\ktype$ to $\gamma$, add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $cs$, and return $?t$ as the type of $e$.
+  For the expression $\sfst~e$ we first recursively infer $e$: $\func{inferExpr}~e = \gamma,\Gamma,\tau,Cs$.
+  We then add two fresh metavariables $?t$ and $?s$ of kind $\ktype$ to $\gamma$, add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $cs$, then return $?t$ as the type of $e$.
 
 \item
-  For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \gamma_p, (\Gamma_p, s) , Cs_p$, $\func{inferProc}~q = \gamma_q, (\Gamma_q, t) , Cs_q$, and $\func{inferExpr}~e = \gamma_e, \Gamma_e, \tau, Cs_e$.
-  We add the metavariables in $\gamma_p$ to the ones in $\gamma_q$ and a new metavariable $?r$ of kind $\ktype$, and refer to the result as $\delta$.
-  We add the constraints in $Cs_p$  to the ones in $Cs_q$ and create new constraints $\eqconstr{\Gamma_p}{\Gamma_q}$ 
+  For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \theta_p, (\Theta_p, s) , Cs_p$, $\func{inferProc}~q = \theta_q, (\Theta_q, t) , Cs_q$, and $\func{inferExpr}~e = \delta_e, \Delta_e, \tau, Cs_e$.
+  We create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$.
+  We take the union of the metavariables in $\gamma$, $\theta_p$, $\theta_q$ and $\delta_e$
+  We take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tprod~s~t}$ and return $\Gamma$ as the inferred context.
 \end{itemize}

### 2021-05-11 paper/paper.tex
Format constraint resolution

@@
 \newcommand{\subst}[2]{#1 ~ \func{\triangleleft} ~ #2}
+\newcommand{\tSubst}[2]{\type{Subst}~#1~#2}
+\newcommand{\interpr}[1]{\func{[\![} #1 \func{]\!]}}
 
@@
 
-\title{Extended Abstract}
-\subtitle{Co-Contextual Type Inference for the Linear \Picalc{} in Agda}
+\title
+[Co-Contextual Type Inference for the Linear \Picalc{} in Agda]
+{Co-Contextual Type Inference \\ for the Linear \Picalc{} \\ in Agda}
+\subtitle{(Extended Abstract)}
 
@@
 
-(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and I am sorry this is not a wee poem.)
+(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and we are sorry this is not a wee poem.)
 
@@
 \end{mathpar}
+\todo{generalisation to an algebra?}
 
@@
 
+\todo{Introduce Subst type}
+
 We encode substitution as a function $\subst{\_}{\_}$ of type $(\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}) \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ that acts on usage annotations and types and typing contexts, and that freely alters the metavariable context.
@@
 
-We use a $\func{[\![\_]\!]}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.
+We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.
 
@@
 $$
+This functions are total: if processes and expressions are untypable, their constraints will simply be unsolvable.
 Let us provide a couple of examples illustrating how this function is defined:
@@
   We take the union of the metavariables in $\gamma$, $\theta_p$, $\theta_q$ and $\delta_e$
-  We take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tprod~s~t}$ and return $\Gamma$ as the inferred context.
+  We take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tsum~s~t}$ and return $\Gamma$ as the inferred context.
 \end{itemize}
@@
 
-Every substitution \(\sigma\) that makes the constraints hold will make
-the process typable under the substituted context. \[
-\begin{aligned}
-\texttt{infer-sound} &: \forall (p : \texttt{Proc}~n). ~ \texttt{infer}~P \equiv \gamma , \Gamma , cs \\
-&\to \forall (\sigma : \texttt{Subst}~\gamma~\delta). ~ [\![ \sigma \triangleleft cs ]\!] \\
-&\to (\sigma \triangleleft \Gamma) \vdash P
-\end{aligned}
-\]
+Every substitution \(\sigma\) that makes the constraints hold will make the process typable under the substituted context.
+\begin{flalign*}
+& \tEq{\func{infer}~p}{\gamma , \Gamma , cs} \to (\sigma : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma}{cs}} \to \tProc{(\subst{\sigma}{\Gamma})}{p} &&
+\end{flalign*}
 
@@
 
-For every context \(\Delta\) that makes the process typable there exists
-a most general substitution \(\sigma\) that will solve the constraints
-and which substitutes in \(\Gamma\) to something more general than
-\(\Delta\). \[
-\begin{aligned}
-\texttt{infer-complete} &: \forall (p : \texttt{Proc}~n). ~ \texttt{infer}~P \equiv \gamma , \Gamma , cs \\
-&\to \forall (\Delta : \texttt{Ctx}~n~\gamma). ~ \Delta \vdash P \\
-&\to \exists \delta. ~ \exists (\sigma : \texttt{Subst}~\gamma~\delta). \\
-&\to [\![ \sigma \triangleleft cs ]\!] \times \Delta \subseteq (\sigma \triangleleft \Gamma)
-\end{aligned}
-\] where \[
-\Delta \subseteq \Gamma \triangleq \exists \delta. ~ \exists (\sigma : \texttt{Subst} \_ \delta). ~ \Delta \equiv (\sigma \triangleleft \Gamma)
-\]
+For every context \(\Delta\) that makes the process typable there exists a most general substitution \(\sigma\) that will solve the constraints and which substitutes in \(\Gamma\) to something more general than \(\Delta\).
+\begin{flalign*}
+& \tEq{\func{infer}~p}{\gamma , \Gamma , cs} \to (\Delta : \tCtx{n}{\gamma}) \to \tProc{\Delta}{p} && \\
+&\to \exists \delta. ~ \exists (\sigma : \texttt{Subst}~\gamma~\delta). \interpr{\subst{\sigma}{cs}} \times (\tProc{\subst{\sigma}{\Gamma}}{p}) \times \Delta \subseteq (\subst{\sigma}{\Gamma}) &&
+\end{flalign*}
+where $$
+\Delta \func{\subseteq} \Gamma \triangleq \exists \delta. ~ \exists (\sigma : \tSubst{\_}{\delta}). ~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}
+$$
 
-\hypertarget{constraint-resolution}{%
-\section{Constraint Resolution}\label{constraint-resolution}}
+\section{Constraint Resolution}\label{constraint-resolution}
 
@@
 \begin{aligned}
-\texttt{solve} &: \texttt{Subst}~\gamma~\delta \times \texttt{List}~(\texttt{Constr}~\delta) \\
-&\to \texttt{Subst}~\gamma~\phi \times \texttt{List}~(\texttt{Constr}~\phi)
+\func{solve} &: \tConstrs{\gamma} \to \tSubst{\gamma}{\delta} \times \tConstrs{\delta}
 \end{aligned}
@@
 
-The simplified constraints are enough to entail the original
-constraints. \[
+The simplified constraints are enough to entail the original constraints.
+\[
 \begin{aligned}
-\texttt{solve-sound}
-&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
-&\to \forall (\sigma_f : \texttt{Subst}~\gamma~\delta) \\
-& \to [\![ \sigma_f \triangleleft cs_2 ]\!] \to [\![ \sigma_f \cdot \sigma_2 \triangleleft cs_1 ]\!]
+& ~ \tEq{\func{solve}~cs_1}{(\sigma, cs_2)} \to (\sigma_f : \tSubst{\gamma}{\delta}) \\
+\to& \interpr{\subst{\sigma_f}{cs_2}} \to \interpr{\subst{\sigma_f \cdot \sigma}{cs_1}}
 \end{aligned}
@@
 
-Any substitution \(\sigma_f\) that makes the original constraints
-\(cs_1\) hold will be a specialisation \(\sigma_f\) of our returned
-substitution \(\sigma_2\). The specialisation \(\sigma_g\) will make the
-returned constraints \(cs_2\) hold. \[
-\begin{aligned}
-\texttt{solve-sound}
-&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
-&\to \forall (\sigma_f : \texttt{Subst}~\gamma~\delta). ~ [\![ \sigma_f \triangleleft cs_1 ]\!] \\
-&\to \exists (\sigma_g : \sigma_f = \sigma_g \cdot \sigma_2) \times [\![ \sigma_g \triangleleft cs_2 ]\!]
-\end{aligned}
-\]
+Any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold will be a specialisation \(\sigma_f\) of our returned substitution \(\sigma_2\).
+The specialisation \(\sigma_g\) will make the returned constraints \(cs_2\) hold.
+\begin{flalign*}
+&\tEq{\func{solve}~cs_1}{(\sigma, cs_2)} \to (\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_1}} && \\
+&\to \exists (\sigma_g : \tSubst{\_}{\delta}). ~ \tEq{\sigma_f}{\sigma_g \cdot \sigma} \times \interpr{\subst{\sigma_g}{cs_2}} &&
+\end{flalign*}
 
@@
 unsolvable because we have not enough information, or is unsatisfiable
-altogether. \[
-\begin{aligned}
-\texttt{solve-sound}
-&: \texttt{solve}~(\sigma_1, cs_1) \equiv (\sigma_2, cs_2) \\
-&\to \forall c \in cs_2. ~ \texttt{IsDeferred}~c \uplus \texttt{IsUnsat}~c
-\end{aligned}
-\] where \[
-\texttt{IsUnsat}~c = \forall \sigma. ~ \neg [\![ \sigma \triangleleft c ]\!]
-\]
+altogether.
+\begin{flalign*}
+&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsDeferred}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
+\end{flalign*}
+where
+\begin{flalign*}
+& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} &&
+\end{flalign*}
 
@@
 
-\[
-\forall c. ~ \texttt{IsDeferred}~c \to \exists \sigma. [\![ \sigma \triangleleft c ]\!]
-\]
+\begin{flalign*}
+& \forall c. ~ \type{IsDeferred}~c \to \exists \sigma. \interpr{\subst{\sigma}{c}} &&
+\end{flalign*}
 

### 2021-05-11 paper/paper.tex
Another go at the introduction

@@
   The linear \picalc{} is a well-known substructural type system for the \picalc{}:
-  it is a common target for the encoding of session-typed languages \cite{DardhaGS17},
-  and it in direct correspondence with classical linear logic \todo{cite}.
+  it is a common target for the encoding of session-typed languages \cite{DardhaGS17}
+  and in direct correspondence with classical linear logic \todo{cite}.
 
-  Type inference for the linear \picalc{} comes with its own challenges:
-  for one, processes themselves are untyped, and thus typing constraints only apply to the typing contexts;
-  moreover, typing constraints on usage annotations \emph{cannot always be eagerly solved in open processes}.
+  For the type checking of the linear \picalc{} to be decidable the newly created channels must have their types synthesized.
+  Inferring these types can prove challenging, as not all their typing constraints are local.
+  Moreover, some typing constraints cannot be eagerly solved and need to be deferred.
 
-  We approach the first challenge using \emph{co-contextual type inference} \cite{ErdwegBKKM15}, where processes are traversed bottom-up and constraints are collected (and when possible eagerly solved) and merged on context splits.
-  We approach the latter challenge by allowing the resolution of some of the constraints to be deferred.
-  We state soundness and completeness theorems that take this deferral of constraint resolution into account.
-  We state a progress (?) theorem that shows that only those constraints that need to be deferred are deferred.
-  \todo{be more precise}
-  We are working on mechanising our work in Agda.
+  Following \cite{Padovani15}, we split this problem in two: first we provide a \emph{co-contextual type inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints and merging them on context splits; then we solve some of these constraints using well-known unification algorithms \cite{McBride03}, while we defer the resolution of those constraints that lack enough information.
+  We provide clear soundness and completeness theorems for both these stages, and show that only those constraints that lack enough information are deferred.
+  The totality of this work is being mechanised in Agda.
 \end{abstract}
@@
 
-We define an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices \autoref{syntax}.
-We define a small expression language within the process syntax to handle composite sum and product types.
-On top of the well-scoped syntax we provide the standard linear type-system using context-splits \autoref{type-system}.
-The type system is indexed by a typing context that admits unsolved metavariables.
-These unsolved metavariables are however well kinded: either types or usage annotations.
+Consider a linear \picalc{} process with a free channel variable $x$ and wherein we create a new channel $y$, use $y$ to transmit, then send $y$ away over $x$ and terminate.
+Unless the payload type of $x$ is determined elsewhere, in this process the usage annotations for the channel $y$ are polymorphic, yet the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
+While some of the constraints on $x$ and $y$ can be eagerly solved, others are non-local and need to be deferred.
 
-\todo{insert example showing why we need deferral}
+We approach this problem by breaking down typing inference into the collection of constraints \autoref{inference} and their (partial) resolution \autoref{constraint-resolution}.
+We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices \autoref{syntax}.
+The syntax for processes contains a small expression syntax that handles composite sum and product types.
+On top of the well-scoped syntax we define a standard linear type system using context-splits \autoref{type-system}.
 
-We split typing inference into constraint collection and constraint solving \autoref{inference}.
-We start by defining constraints on types and usage annotations \autoref{constraints}
-We collect constraints bottom-up, creating extra \emph{merging} constraints on context splits.
-This allows type-inference to be easily parallelisable.
-We postulate soundness and completeness theorems that allow for constraint solving to be deferred \autoref{inference-theorems}.
+Constraints are defined on types and usage annotations alike \autoref{constraints}.
+We use a common metavariable context for both types and usage annotations.
+The metavariable context is kinded, ans substitution is defined so that it respects kindedness.
+In \autoref{inference} constraints are collected bottom-up, making type-inference easily parallelisable.
+Context splits are solved by creating a fresh extra context and constraints that \emph{merge} the three.
+We prove a soundness theorem and postulate a completeness theorem, both allowing for constraint solving to be deferred \autoref{inference-theorems}.
+
+Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of \cite{McBride03}.
+Constraints of the form $\eqconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
+We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred.
 

### 2021-05-11 paper/paper.tex
Rewrite the intro

@@
 Consider a linear \picalc{} process with a free channel variable $x$ and wherein we create a new channel $y$, use $y$ to transmit, then send $y$ away over $x$ and terminate.
-Unless the payload type of $x$ is determined elsewhere, in this process the usage annotations for the channel $y$ are polymorphic, yet the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
-While some of the constraints on $x$ and $y$ can be eagerly solved, others are non-local and need to be deferred.
+Unless the payload type of $x$ is determined elsewhere, the usage annotations for the channel $y$ are polymorphic, and the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
+This lower bound is a constraint that cannot be solved locally: it needs to be deferred until it is further constrained.
 
-We approach this problem by breaking down typing inference into the collection of constraints \autoref{inference} and their (partial) resolution \autoref{constraint-resolution}.
-We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices \autoref{syntax}.
-The syntax for processes contains a small expression syntax that handles composite sum and product types.
-On top of the well-scoped syntax we define a standard linear type system using context-splits \autoref{type-system}.
+We break down typing inference into the collection of constraints (\autoref{inference}) and their (partial) resolution (\autoref{constraint-resolution}).
+We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices (\autoref{syntax}).
+The syntax for processes contains a small expression language that handles composite sum and product types.
+On top of the well-scoped syntax we define a standard linear type system that uses context-splits (\autoref{type-system}).
 
-Constraints are defined on types and usage annotations alike \autoref{constraints}.
-We use a common metavariable context for both types and usage annotations.
-The metavariable context is kinded, ans substitution is defined so that it respects kindedness.
-In \autoref{inference} constraints are collected bottom-up, making type-inference easily parallelisable.
+Types and usage annotations contain kinded metavariables (holes) part of a common metavariable context.
+Substitution respects kindedness, and constraints are uniformly defined over both types and usage annotations.
+Constraints are collected bottom-up (\autoref{inference}), which makes type-inference easily parallelisable.
 Context splits are solved by creating a fresh extra context and constraints that \emph{merge} the three.
-We prove a soundness theorem and postulate a completeness theorem, both allowing for constraint solving to be deferred \autoref{inference-theorems}.
+We prove a soundness theorem and postulate a completeness theorem, both use deferred constraints (\autoref{inference-theorems}).
 
 Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of \cite{McBride03}.
-Constraints of the form $\eqconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
-We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred.
+Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
+We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution-theorems}).
 
@@
 
-Nothing prevents us from returning the original constraints as output.
-We therefore promise that every constraint we return is either currently
-unsolvable because we have not enough information, or is unsatisfiable
+To keep us from returning the original constraints as output we show that every constraint we return is either unsatisfiable or not uniquely constrained and has to be deferred.
 altogether.
@@
 \begin{flalign*}
-& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} &&
-\end{flalign*}
-
-\paragraph{Instantiation.}
-
-\begin{flalign*}
-& \forall c. ~ \type{IsDeferred}~c \to \exists \sigma. \interpr{\subst{\sigma}{c}} &&
+& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} && \\
+& \type{IsNotUnique}~c \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 &&
 \end{flalign*}

### 2021-05-11 paper/paper.tex
Add recursion

@@
 \newcommand{\scase}{\constr{case}}
+\newcommand{\srec}{\constr{rec}}
 
@@
   |&~ \scomp~p_n~q_n \\
-  |&~ \scase~e_n~p_{1+n}~q_{1+n}
+  |&~ \scase~e_n~p_{1+n}~q_{1+n} \\
 \end{aligned}
 \]
-\todo{add recursion}
 \todo{add type assertions}
@@
 \end{mathpar}
-We will henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
+We henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
 
@@
              { \tVar{\Gamma,t}{\fzero}{t} }
-             \rulename{zero}
+             % \rulename{zero}
 
@@
              { \tVar{\Gamma,s}{\fsuc~i}{t} }
-             \rulename{suc}
+             % \rulename{suc}
 
@@
 
-  \inferrule { \tVar{\Gamma}{i}{t} } { \tExpr{\Gamma}{\svar~i}{t} } \rulename{var}
+  \inferrule { \tVar{\Gamma}{i}{t} } { \tExpr{\Gamma}{\svar~i}{t} }
+             % \rulename{var}
 
-  \inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} } \rulename{unit}
+  \inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} }
+             % \rulename{unit}
 
-  \inferrule { \tExpr{\Gamma}{e}{\tprod~t~s} \\ \tun{s}} { \tExpr{\Gamma}{\sfst~e}{t}} \rulename{fst}
+  \inferrule { \tExpr{\Gamma}{e}{\tprod~t~s} \\ \tun{s}} {\tExpr{\Gamma}{\sfst~e}{t}}
+             % \rulename{fst}
 
-  \inferrule { \tExpr{\Gamma}{e}{t} } { \tExpr{\Gamma}{\sinl~e}{\tsum~t~s}} \rulename{inl}
+  \inferrule { \tExpr{\Gamma}{e}{t} } { \tExpr{\Gamma}{\sinl~e}{\tsum~t~s}}
+             % \rulename{inl}
 
@@
              { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
-             \rulename{pair}
+             %\rulename{pair}
 
@@
 
-  \inferrule { \tun{\Gamma} } { \tProc{\Gamma}{\send} } \rulename{end}
+  \inferrule { \tun{\Gamma} } { \tProc{\Gamma}{\send} }
+             %\rulename{end}
+
+  \inferrule { \tProc{\Gamma}{p} \\ \tun{\Gamma} }
+             { \tProc{\Gamma}{\srec~p} }
+             %\rulename{rec}
 
   \inferrule { t : \ttype{\gamma} \\ \tProc{\Gamma , t}{p} } { \tProc{\Gamma}{\snew~p} }
-             \rulename{new}
+             %\rulename{new}
 
@@
              { \tProc{\Gamma}{\srecv~e~p} }
-             \rulename{recv}
+             %\rulename{recv}
 
@@
              { \tProc{\Gamma}{\scase~e~p~q} }
-             \rulename{case}
+             %\rulename{case}
 \end{mathpar}
@@
 
-Constraints are of type $\tConstr{\gamma}$ and are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, both usage annotations and types.
+Constraints of type $\tConstr{\gamma}$ are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, both usage annotations and types.
 They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
@@
 
-\paragraph{Constraint satisfaction}
+\paragraph{Constraint Satisfaction}
 
@@
 
-\paragraph{Inference soundness.}
+\paragraph{Inference Soundness}
 
@@
 
-\paragraph{Inference completeness.}
+\paragraph{Inference Completeness}
 
@@
 
-\section{Constraint Resolution}\label{constraint-resolution}
+\paragraph{Constraint Resolution}
 
+\todo{talk about unification}
 \cite{McBride03}
-
-\paragraph{Constraint Resolution.}
+\label{constraint-resolution}
 
@@
 
-\paragraph{Soundness.}
+\paragraph{Resolution Soundness}
 
@@
 
-\paragraph{Completeness.}
+\paragraph{Resolution Completeness}
 
@@
 
-\paragraph{Progress.}
+\paragraph{Resolution Progress}
 

### 2021-05-12 paper/paper.tex
Another pass to the intro

@@
 \usepackage{mathpartir}
-\mprset {sep=1em} % Horizontal space between premises
+\mprset {sep=0.5em} % Horizontal space between premises
 
@@
 \begin{abstract}
-  The linear \picalc{} is a well-known substructural type system for the \picalc{}:
-  it is a common target for the encoding of session-typed languages \cite{DardhaGS17}
-  and in direct correspondence with classical linear logic \todo{cite}.
+  The linear \picalc{} is a well-known substructural type system for the \picalc{} and a common target for the encoding of session-typed languages \cite{DardhaGS17}.
 
-  For the type checking of the linear \picalc{} to be decidable the newly created channels must have their types synthesized.
-  Inferring these types can prove challenging, as not all their typing constraints are local.
-  Moreover, some typing constraints cannot be eagerly solved and need to be deferred.
+  To make the type checking of the linear \picalc{} decidable newly created channels must have their types synthesized.
+  Inferring these types can prove challenging: not all their typing constraints are local and some cannot be eagerly solved and need to be deferred.
 
-  Following \cite{Padovani15}, we split this problem in two: first we provide a \emph{co-contextual type inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints and merging them on context splits; then we solve some of these constraints using well-known unification algorithms \cite{McBride03}, while we defer the resolution of those constraints that lack enough information.
-  We provide clear soundness and completeness theorems for both these stages, and show that only those constraints that lack enough information are deferred.
+  Following \cite{Padovani15}, we first provide a \emph{co-contextual type inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints; then we solve some of these constraints using well-known unification algorithms \cite{McBride03}, and defer the ones that are unsatisfiable or not uniquely determined.
+  We propose clear soundness and completeness theorems for both these stages and a progress theorem that ensures that only unsatisfiable and non-uniquely determined constraints are deferred.
   The totality of this work is being mechanised in Agda.
@@
 Unless the payload type of $x$ is determined elsewhere, the usage annotations for the channel $y$ are polymorphic, and the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
-This lower bound is a constraint that cannot be solved locally: it needs to be deferred until it is further constrained.
+This lower bound is a constraint that cannot be solved locally: it needs to be deferred until it is further constrained elsewhere.
 
-We break down typing inference into the collection of constraints (\autoref{inference}) and their (partial) resolution (\autoref{constraint-resolution}).
-We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices (\autoref{syntax}).
-The syntax for processes contains a small expression language that handles composite sum and product types.
-On top of the well-scoped syntax we define a standard linear type system that uses context-splits (\autoref{type-system}).
+We use co-contextual type inference \cite{ErdwegBKKM15} to traverse processes bottom-up and infer a context of \emph{requirements} for any given process.
+Collecting constraints bottom-up makes type-inference easily parallelisable.
+This process is broken down into the collection of constraints (\autoref{inference}) and their (partial) resolution (\autoref{constraint-resolution}).
+We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices (\autoref{syntax}) --- the syntax for processes contains a small expression language that handles composite sum and product types.
+We define a standard linear type system on top of the syntax using context-splits (\autoref{type-system}).
 
-Types and usage annotations contain kinded metavariables (holes) part of a common metavariable context.
+Both types and usage annotations contain metavariables (holes).
+We keep a common context of metavariables for both types and usage annotations and use kinds to distinguish them.
 Substitution respects kindedness, and constraints are uniformly defined over both types and usage annotations.
-Constraints are collected bottom-up (\autoref{inference}), which makes type-inference easily parallelisable.
-Context splits are solved by creating a fresh extra context and constraints that \emph{merge} the three.
-We prove a soundness theorem and postulate a completeness theorem, both use deferred constraints (\autoref{inference-theorems}).
+Context splits are solved by creating an extra fresh context together with constraints that bind the three.
+We prove that the collection of constraints is sound and postulate a completeness theorem (\autoref{inference-theorems}).
 
-Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of \cite{McBride03}.
-Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, when they are either admitted as valid or deferred.
-We postulate soundness and completeness theorems and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution-theorems}).
+Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of McBride's unification by structural recursion \cite{McBride03}.
+Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they are either admitted as valid (and result in a substitution), not uniquely determined, or unsatisfiable.
+We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution-theorems}).
 
-(Variables are black, $\type{\textrm{types are blue}}$, $\constr{\textrm{constructors are olive green}}$, $\func{\textrm{functions are gray}}$, and we are sorry this is not a wee poem.)
+(Variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are olive green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a wee poem.)
 
-\section{Syntax}\label{syntax}
+\section{Type System}
+
+\paragraph{Syntax}
+\label{syntax}
 
 Expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$, variable references $i_n$ are of type $\Fin{n}$.
+\todo{add type assertions}
 \[
@@
 \]
-\todo{add type assertions}
 
-\section{Type System}\label{type-system}
+\paragraph{Types}
+\label{types}
 
@@
 
+\paragraph{Typing Judgments}
+\label{typing-judgments}
 We can finally define our typing judgment for variables, expressions and processes -- some are omitted for brevity.
@@
 
-\todo{Introduce Subst type}
-
-We encode substitution as a function $\subst{\_}{\_}$ of type $(\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}) \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ that acts on usage annotations and types and typing contexts, and that freely alters the metavariable context.
-We define substitutions on constraints pointwise on their arguments.
+A substitution $\tSubst{\gamma}{\delta}$ maps every metavariable of kind $k$ in $\gamma$ to a term of kind $k$ in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$
+The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
+Substitutions on constraints are defined pointwise on their arguments.
 
@@
 
-We encode type inference as a function taking processes with $n$ free variables and returning a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$:
-$$
-\func{inferProc} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma}
-$$
-We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression.
-$$
-\func{inferExpr} : \sExpr{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \ttype{\gamma} \times \tConstrs{\gamma}
-$$
-This functions are total: if processes and expressions are untypable, their constraints will simply be unsolvable.
+We encode type inference as a function taking processes with $n$ free variables and returning a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$.
+We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression:
+\begin{flalign*}
+& \func{inferProc} : \sProc{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma} && \\
+& \func{inferExpr} : \sExpr{n} \to ~ \exists \gamma. ~ \tCtx{n}{\gamma} \times \tConstrs{\gamma} \times \ttype{\gamma} &&
+\end{flalign*}
+These functions are total: if processes and expressions are untypable, their constraints will simply be unsolvable.
 Let us provide a couple of examples illustrating how this function is defined:
@@
 \item
-  For the expression $\sfst~e$ we first recursively infer $e$: $\func{inferExpr}~e = \gamma,\Gamma,\tau,Cs$.
+  For the expression $\sfst~e$ we first recursively infer $e$: $\func{inferExpr}~e = \gamma,\Gamma,Cs,\tau$.
   We then add two fresh metavariables $?t$ and $?s$ of kind $\ktype$ to $\gamma$, add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $cs$, then return $?t$ as the type of $e$.
@@
 \item
-  For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \theta_p, (\Theta_p, s) , Cs_p$, $\func{inferProc}~q = \theta_q, (\Theta_q, t) , Cs_q$, and $\func{inferExpr}~e = \delta_e, \Delta_e, \tau, Cs_e$.
+  For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \theta_p, (\Theta_p, s) , Cs_p$, $\func{inferProc}~q = \theta_q, (\Theta_q, t) , Cs_q$, and $\func{inferExpr}~e = \delta_e, \Delta_e, Cs_e, \tau$.
   We create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$.
@@
 \begin{flalign*}
-&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsDeferred}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
+&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsNotUnique}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
 \end{flalign*}
@@
 \begin{flalign*}
-& \type{IsUnsat}~c \triangleq \forall \sigma. ~ \neg \interpr{\subst{\sigma}{c}} && \\
-& \type{IsNotUnique}~c \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 &&
+& \type{IsNotUnique}~c && \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 && \\
+& \type{IsUnsat}~c && \triangleq \neg \exists \sigma. ~ \interpr{\subst{\sigma}{c}} &&
 \end{flalign*}

### 2021-05-12 paper/paper.tex
Another pass

@@
 Collecting constraints bottom-up makes type-inference easily parallelisable.
-This process is broken down into the collection of constraints (\autoref{inference}) and their (partial) resolution (\autoref{constraint-resolution}).
-We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices (\autoref{syntax}) --- the syntax for processes contains a small expression language that handles composite sum and product types.
-We define a standard linear type system on top of the syntax using context-splits (\autoref{type-system}).
+This process is broken down into the collection of constraints (\autoref{constraint-collection}) and their (partial) resolution (\autoref{constraint-resolution}).
+We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices --- the syntax for processes contains a small expression language that handles composite sum and product types.
+We define a standard linear type system on top of the syntax using context-splits.
 
@@
 Context splits are solved by creating an extra fresh context together with constraints that bind the three.
-We prove that the collection of constraints is sound and postulate a completeness theorem (\autoref{inference-theorems}).
+We prove that the collection of constraints is sound and postulate a completeness theorem (\autoref{constraint-collection}).
 
@@
 Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they are either admitted as valid (and result in a substitution), not uniquely determined, or unsatisfiable.
-We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution-theorems}).
+We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution}).
 
@@
 
+We define a standard syntax and type system for the linear \picalc{}.
+The only non-standard feature is that types allow for metavariables (holes) within them.
+
 \paragraph{Syntax}
@@
 
-Expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$, variable references $i_n$ are of type $\Fin{n}$.
+We define a standard syntax for the \picalc{} using type-level de Bruijn indices.
+Variable references $i_n$ are of type $\Fin{n}$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
 \todo{add type assertions}
@@
   p_n ~ q_n  :=
-  &~ \send ~|~  \snew~p_{1+n} \\
+  &~ \send ~|~  \srec~p_n ~|~ \snew~p_{1+n} \\
   |&~ \srecv~e_n~p_{1+n} \\
@@
 Both types and usage annotations allow for \emph{holes} or \emph{metavariables}.
-We use the same set of metavariables for both types and usage annotations -- this makes their handling more uniform.
-We use a context of kinds \(\gamma\) to keep track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$: a choice of kind \(k\) in a kinding context \(\gamma\) is denoted as $\tvar{\gamma}{k}$.
-
-We define well kinded types and usage annotations simultaneously so that we are able to provide a uniform unification and context merge algorithm later on (note that we define $\tsum$ and $\tprod$ and $\tzero$, $\tone$ and $\tomega$ in one go to save space):
+We use a common set of metavariables for both types and usage annotations -- this makes their handling uniform.
+A context of kinds \(\gamma\) keeps track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$.
+We refer to a metavariable of kind \(k\) in a kinding context \(\gamma\) as $\tvar{\gamma}{k}$.
+We can now define usage annotations and types in one go (note that we define $\tsum$ and $\tprod$ together for brevity):
 \begin{mathpar}
@@
 
-  \inferrule { } {\tzero ~|~ \tone ~|~ \tomega : \tkind{\gamma}{\kusage}}
+  \inferrule { } {\tzero : \tkind{\gamma}{\kusage}}
+
+  \inferrule { } {\tone : \tkind{\gamma}{\kusage}}
+
+  \inferrule { } {\tomega : \tkind{\gamma}{\kusage}}
 \end{mathpar}
-We henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
+For convenience we henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
 
+\paragraph{Context Splits}
+\label{types}
 A context $\Gamma$ of type $\tCtx{n}{\gamma}$ is a list of $\ttype{\gamma}$ of size $n$.
 We define context splits $\tSplit{\Gamma}{\Delta}{\Theta}$ pointwise on types.
-Splits on types are defined pointwise recursively on usage annotations -- note however that only the usage annotations on the top level of a channel (that is, not within its payload) are split.
+Splits on types are defined recursively pointwise on usage annotations -- note however that only the usage annotations on the top level of a channel (that is, not within its payload) are split.
 Splits on usage annotations are defined as follows -- note that they are not necessarily unique:
@@
 \end{mathpar}
-\todo{generalisation to an algebra?}
-
-A usage annotation $x$, type $t$ or context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $x$ and $t$) \cite{Padovani15}.
+Following \cite{Padovani15}, we say a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ --- (similarly $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$, respectively).
 
@@
 
-Co-contextual type inference traverses processes bottom up collecting constraints.
-Limiting the information flow to being bottom up has the advantage of making inference subproblems independent and thus easily parallelisable.
+Co-contextual type inference traverses processes bottom-up collecting constraints.
+This has the advantage of making inference subproblems independent and thus easily parallelisable.
 For the shared \picalc{}, inference constraints amount to unification problems that can always be eagerly solved.
-The linear \picalc{} however requires that some constraints are deferred to a later stage, when more information is available.
+The linear \picalc{} however requires that some constraints are deferred to a later stage, when more information becomes available elsewhere.
 
-Let us suppose we can safely instantiate to $\tzero$ every usage annotation that is unrestricted.
 Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we send the unit over $a$, then send $a$ over $x$ and terminate.
@@
 Since we send a unit over $a$, $a$ must be of some type $s$ such that $\tSplit{s}{\tchan~\tzero~\tone~\tunit}{t}$.
-From this we can infer that both $s$ and $t$ must be a channel type and thus $\tSplit{s_i}{\tzero}{t_i}$ and $\tSplit{s_o}{\tone}{t_o}$ for some $s_i$, $t_i$, $s_o$ and $t_o$, but we cannot infer their usage annotations: all we can deduce is their lower bounds $\tzero$ and $\tone$.
-We must defer the resolution of these constraints until more information becomes available about the context of this open process.
+From this we can infer that both $s$ and $t$ must be a channel type and thus $\tSplit{s_i}{\tzero}{t_i}$ and $\tSplit{s_o}{\tone}{t_o}$ for some $s_i$, $t_i$, $s_o$ and $t_o$.
+We cannot however infer these usage annotations: all we can deduce is their lower bounds $\tzero$ and $\tone$.
+The resolution of these constraints must be deferred until more information becomes available in the context of this open process.
 
-We present two kinds of constraints, showcase how to encode inference problems for some of the typing rules in \autoref{type-system}, and state the lemmas that will allow us to show that our approach to inference is both sound and complete.
+In \autoref{constraint-collection} we introduce typing constraints, define substitution and constraint satisfaction, and provide a typing inference algorithm.
+We show that satisfying the constraints generated by typing inference is enough to make a process typable (soundness), and postulate that for any typable process typing inference will find the most general set of constraints that makes the process typable (completeness).
+
+In \autoref{constraint-resolution} we solve equality constraints by unification and context split constraints by decomposition.
+Constraints have either no solution, a unique solution, or multiple possible solutions.
+(Committing to a non-unique solution can prove to be the wrong choice in an open process, and thus we avoid doing so.
+Closed processes can have non-unique solutions safely instantiated.)
+Solving a set of constraints results in a set of substitutions and a set of simplified constraints where those substitutions have already been performed.
+We postulate that satisfying the simplified constraints amounts to satisfying the original constraints after substitution (soundness), and that for every substitution that solves the original constraints we can find a most general substitution that will solve the simplified constraints (completeness).
+Additionally, we postulate that the simplified set of constraints only contains unsatisfiable or non-uniquely satisfiable constraints (progress).
+
+\subsection{Constraint Collection}
+\label{constraint-collection}
 
@@
 They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
-Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter \emph{sometimes} cannot, and must be deferred.
+Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter can \emph{sometimes} not, and must be deferred.
 We use $\tConstrs{\gamma}$ to refer to lists of constraints of type $\tConstr{\gamma}$.
@@
 
-A substitution $\tSubst{\gamma}{\delta}$ maps every metavariable of kind $k$ in $\gamma$ to a term of kind $k$ in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$
+A substitution $\tSubst{\gamma}{\delta}$ maps usage annotation and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
 The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
@@
 
-We encode type inference as a function taking processes with $n$ free variables and returning a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$.
+We encode type inference as a function that takes processes with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$.
 We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression:
@@
 Let us provide a couple of examples illustrating how this function is defined:
-\begin{itemize}
-\item
+
 For the expression $\sfst~e$ we first recursively infer $e$: $\func{inferExpr}~e = \gamma,\Gamma,Cs,\tau$.
@@
 
-\item
 For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \theta_p, (\Theta_p, s) , Cs_p$, $\func{inferProc}~q = \theta_q, (\Theta_q, t) , Cs_q$, and $\func{inferExpr}~e = \delta_e, \Delta_e, Cs_e, \tau$.
@@
 We take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tsum~s~t}$ and return $\Gamma$ as the inferred context.
-\end{itemize}
 
@@
 
-Every substitution \(\sigma\) that makes the constraints hold will make the process typable under the substituted context.
+Every substitution \(\sigma\) that makes the constraints hold will make the process typable under the substituted context. Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$ we have that:
 \begin{flalign*}
-& \tEq{\func{infer}~p}{\gamma , \Gamma , cs} \to (\sigma : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma}{cs}} \to \tProc{(\subst{\sigma}{\Gamma})}{p} &&
+& (\sigma : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma}{cs}} \to \tProc{(\subst{\sigma}{\Gamma})}{p} &&
 \end{flalign*}
@@
 For every context \(\Delta\) that makes the process typable there exists a most general substitution \(\sigma\) that will solve the constraints and which substitutes in \(\Gamma\) to something more general than \(\Delta\).
+We define $\Delta ~ \func{\subseteq} ~ \Gamma$ as $\exists \delta. ~ \exists (\sigma : \tSubst{\_}{\delta}). ~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}$.
+Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$ we have that:
 \begin{flalign*}
-& \tEq{\func{infer}~p}{\gamma , \Gamma , cs} \to (\Delta : \tCtx{n}{\gamma}) \to \tProc{\Delta}{p} && \\
-&\to \exists \delta. ~ \exists (\sigma : \texttt{Subst}~\gamma~\delta). \interpr{\subst{\sigma}{cs}} \times (\tProc{\subst{\sigma}{\Gamma}}{p}) \times \Delta \subseteq (\subst{\sigma}{\Gamma}) &&
+  &(\Delta : \tCtx{n}{\gamma}) \to \tProc{\Delta}{p} \to && \\
+  & \exists \delta. ~ \exists (\sigma : \tSubst{\gamma}{\delta}). ~ \interpr{\subst{\sigma}{cs}} \times \Delta \subseteq (\subst{\sigma}{\Gamma}) &&
 \end{flalign*}
-where $$
-\Delta \func{\subseteq} \Gamma \triangleq \exists \delta. ~ \exists (\sigma : \tSubst{\_}{\delta}). ~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}
-$$
 
-\paragraph{Constraint Resolution}
-
-\todo{talk about unification}
-\cite{McBride03}
+\subsection{Constraint Resolution}
 \label{constraint-resolution}
 
-Given a set of accumulated substitutions and some constraints, we will
-return (hopefully) some further substitutions, and a set of leftover
-constraints to which the substitutions have been applied. These leftover
-constraints can either not be simplified or are unsatisfiable. \[
+Solving a set of constraints results in a set of substitutions and an unsolved set of simplified constraints where those substitutions have already been performed.
+Simplified constraints have either no solution, or multiple possible solutions.
+Equality constraints are solved by unification following McBride \cite{McBride03}, and are either unsatisfiable or uniquely satisfied --- and thus resulting in a substitution.
+Context split constraints are solved recursively until a base case is reached.
+$$
 \begin{aligned}
@@
 \end{aligned}
-\]
+$$
 
@@
 
-The simplified constraints are enough to entail the original constraints.
+Satisfying the simplified constraints satisfies the original constraints once substitutions are applied: $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$ we have that:
 \[
 \begin{aligned}
-& ~ \tEq{\func{solve}~cs_1}{(\sigma, cs_2)} \to (\sigma_f : \tSubst{\gamma}{\delta}) \\
-\to& \interpr{\subst{\sigma_f}{cs_2}} \to \interpr{\subst{\sigma_f \cdot \sigma}{cs_1}}
+& (\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_2}} \to \interpr{\subst{\sigma_f \cdot \sigma}{cs_1}}
 \end{aligned}
@@
 
-Any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold will be a specialisation \(\sigma_f\) of our returned substitution \(\sigma_2\).
-The specialisation \(\sigma_g\) will make the returned constraints \(cs_2\) hold.
+Any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold will be a specialisation \(\sigma_g\) of the returned substitution \(\sigma\).
+The specialisation \(\sigma_g\) makes the returned constraints \(cs_2\) hold.
+Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$ we have that:
 \begin{flalign*}
-&\tEq{\func{solve}~cs_1}{(\sigma, cs_2)} \to (\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_1}} && \\
-&\to \exists (\sigma_g : \tSubst{\_}{\delta}). ~ \tEq{\sigma_f}{\sigma_g \cdot \sigma} \times \interpr{\subst{\sigma_g}{cs_2}} &&
+&(\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_1}} \to && \\
+&\exists (\sigma_g : \tSubst{\_}{\delta}). ~ \tEq{\sigma_f}{\sigma_g \cdot \sigma} \times \interpr{\subst{\sigma_g}{cs_2}} &&
 \end{flalign*}
@@
 
-To keep us from returning the original constraints as output we show that every constraint we return is either unsatisfiable or not uniquely constrained and has to be deferred.
-altogether.
+To keep us from returning the original constraints as output we postulate that every constraint we return is either unsatisfiable or not uniquely satisfied.
+Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$ we have that:
 \begin{flalign*}
-&\func{solve}~cs_1 \equiv (\sigma, cs_2) \to \forall c \in cs_2. ~ \type{IsNotUnique}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
+&\forall c \to c \in cs_2 \to \type{IsNotUnique}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
 \end{flalign*}

### 2021-05-12 paper/paper.tex
Change the title to two lines
Shorten (sub)section autoref names
Add section with contributions

@@
 \usepackage{todonotes}
+
+\renewcommand*{\sectionautorefname}{\S\!\!\,}
+\renewcommand*{\subsectionautorefname}{\S\!\!\,}
 
@@
 [Co-Contextual Type Inference for the Linear \Picalc{} in Agda]
-{Co-Contextual Type Inference \\ for the Linear \Picalc{} \\ in Agda}
+{Co-Contextual Type Inference \\ for the Linear \Picalc{} in Agda}
 \subtitle{(Extended Abstract)}
@@
 
-(Variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are olive green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a wee poem.)
+\paragraph{Contributions}
+\begin{itemize}
+  \item We break down typing inference into constraint collection and constraint resolution and state clear soundness and completeness theorems for both phases.
+  \item We are working on mechanising the proofs for these theorems in Agda.
+We have already shown the soundness of constraint collection and the soundness of equality constraint resolution (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms).
+\end{itemize}
+
+(Variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
 

### 2021-05-12 paper/paper.tex
Contributions using textbf

@@
 Context splits are solved by creating an extra fresh context together with constraints that bind the three.
-We prove that the collection of constraints is sound and postulate a completeness theorem (\autoref{constraint-collection}).
+\textbf{We prove that the collection of constraints is sound and postulate a completeness theorem} (\autoref{constraint-collection}).
 
@@
 Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they are either admitted as valid (and result in a substitution), not uniquely determined, or unsatisfiable.
-We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution}).
+\textbf{We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred} (\autoref{constraint-resolution}).
 
-\paragraph{Contributions}
-\begin{itemize}
-  \item We break down typing inference into constraint collection and constraint resolution and state clear soundness and completeness theorems for both phases.
-  \item We are working on mechanising the proofs for these theorems in Agda.
-We have already shown the soundness of constraint collection and the soundness of equality constraint resolution (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms).
-\end{itemize}
+\textbf{The totality of this work is being mechanised in Agda.}
+We have shown the soundness of constraint collection and the soundness of equality constraint resolution (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms).
+The remainder proofs are still in progress.
 

### 2021-05-20 paper/paper.tex
Another pass, mostly to the intro

--- 2021-05-12 paper/paper.tex
@@
 
+\usepackage[inline]{enumitem}
 \usepackage{todonotes}
+
+\renewcommand*{\sectionautorefname}{\S\!\!\,}
+\renewcommand*{\subsectionautorefname}{\S\!\!\,}
 
@@
 \title
-[Co-Contextual Type Inference for the Linear \Picalc{} in Agda]
-{Co-Contextual Type Inference \\ for the Linear \Picalc{} \\ in Agda}
+[Co-Contextual Typing Inference for the Linear \Picalc{} in Agda]
+{Co-Contextual Typing Inference \\ for the Linear \Picalc{} in Agda}
 \subtitle{(Extended Abstract)}
@@
 \begin{abstract}
-  The linear \picalc{} is a well-known substructural type system for the \picalc{} and a common target for the encoding of session-typed languages \cite{DardhaGS17}.
+  To type check a \picalc{} with linear and shared types without relying on type annotations, newly created channels must have their types inferred.
+  However, processes can partially use a channel and then send it away: in this case, the receiving process might further constraint its inferred type.
+  This results in non-local constraints that do not have a most general unifier but have to be kept around.
 
-  To make the type checking of the linear \picalc{} decidable newly created channels must have their types synthesized.
-  Inferring these types can prove challenging: not all their typing constraints are local and some cannot be eagerly solved and need to be deferred.
-
-  Following \cite{Padovani15}, we first provide a \emph{co-contextual type inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints; then we solve some of these constraints using well-known unification algorithms \cite{McBride03}, and defer the ones that are unsatisfiable or not uniquely determined.
-  We propose clear soundness and completeness theorems for both these stages and a progress theorem that ensures that only unsatisfiable and non-uniquely determined constraints are deferred.
+  Following \cite{Padovani15}, we provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints.
+  We then solve (some of) these constraints using well-known unification algorithms \cite{McBride03} to find the most general unifier, and defer those that do not have one.
+  We state clear soundness and completeness theorems for both these stages and a progress theorem that ensures that only constraints without a most general unifier are deferred.
   The totality of this work is being mechanised in Agda.
@@
 
-Consider a linear \picalc{} process with a free channel variable $x$ and wherein we create a new channel $y$, use $y$ to transmit, then send $y$ away over $x$ and terminate.
-Unless the payload type of $x$ is determined elsewhere, the usage annotations for the channel $y$ are polymorphic, and the output capability on $y$ requires a lower bound of multiplicity $\tone$ (since the process uses it to transmit).
-This lower bound is a constraint that cannot be solved locally: it needs to be deferred until it is further constrained elsewhere.
+The \picalc{} \todo{cite} models concurrent processing by boiling down process interaction to the transmission of data over communication channels --- where channels too are sent as payload.
+The linear \picalc{} \todo{cite} introduces a type system that ensures that every channel is used \emph{exactly} once.
+This restriction ensures communication privacy, communication safety, and the absence of race conditions.
+We generalise and admit a \picalc{} where the input and output multiplicities of a channel are either $0$ (cannot be used), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
 
-We use co-contextual type inference \cite{ErdwegBKKM15} to traverse processes bottom-up and infer a context of \emph{requirements} for any given process.
-Collecting constraints bottom-up makes type-inference easily parallelisable.
-This process is broken down into the collection of constraints (\autoref{constraint-collection}) and their (partial) resolution (\autoref{constraint-resolution}).
-We start defining an untyped but well scoped \picalc{} using type-level de Bruijn \cite{deBruijn72} indices --- the syntax for processes contains a small expression language that handles composite sum and product types.
-We define a standard linear type system on top of the syntax using context-splits.
+To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within the process.
+To do so we can either ask the user for type annotations \todo{cite}, or try to synthesize these types by looking at how the channels are used.
+The latter approach involves dealing with non-local constraints: a channel can be partially used and then sent away --- the process that receives it might further constrain its type.
+This means that only those constraints with a most general solution can safely be eagerly solved, otherwise we risk over-constraining down the line.
+We use co-contextual typing inference \cite{ErdwegBKKM15} to infer a typing context and gather constraints as we traverse processes bottom-up.
+Typing constraints with a most general unifier are eagerly solved and result in substitutions on the typing context, the remaining typing constraints is carried along.
+Collecting constraints bottom-up has the additional advantage of making typing inference easily parallelisable.
+Empowered with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$.
 
-Both types and usage annotations contain metavariables (holes).
-We keep a common context of metavariables for both types and usage annotations and use kinds to distinguish them.
-Substitution respects kindedness, and constraints are uniformly defined over both types and usage annotations.
-Context splits are solved by creating an extra fresh context together with constraints that bind the three.
-We prove that the collection of constraints is sound and postulate a completeness theorem (\autoref{constraint-collection}).
+To the best of our knowledge, this problem is only treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.
+However, his work does not appear to be mechanised, and metatheoretical properties like soundness and completeness are only informally addressed.
+With this work, we aim to:
+\begin{itemize}
+  \item State and prove clear soundness and completeness theorems for both constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}).
+  \item Mechanise in Agda the totality of this work.
+\end{itemize}
 
-Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of McBride's unification by structural recursion \cite{McBride03}.
-Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they are either admitted as valid (and result in a substitution), not uniquely determined, or unsatisfiable.
-We postulate soundness and completeness theorems for constraint solving and show that only those constraints without a unique solution are deferred (\autoref{constraint-resolution}).
+We start defining an untyped but well scoped \picalc{} using type-level de Bruijn indices \cite{deBruijn72} and embed a small expression language that handles composite sum and product types.
+On top, we define a standard type system with linear and shared types using context-splits.
+We then provide an overview of how type inference works, show how constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}) are defined, and state that both phases are sound and complete with regards to the previously defined type system.
 
-(Variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are olive green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a wee poem.)
+We have proven that constraint collection is sound, and that equality constraint resolution is sound (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms).
+The remainder proofs are still in progress.
+
+(Note on notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
 
@@
 We define a standard syntax and type system for the linear \picalc{}.
-The only non-standard feature is that types allow for metavariables (holes) within them.
+The only non-standard feature is that types allow for \emph{metavariables} (\emph{holes}) within them.
 
@@
 
-Both types and usage annotations allow for \emph{holes} or \emph{metavariables}.
+Both types and usage annotations contain \emph{metavariables} (\emph{holes}).
 We use a common set of metavariables for both types and usage annotations -- this makes their handling uniform.
@@
 This has the advantage of making inference subproblems independent and thus easily parallelisable.
-For the shared \picalc{}, inference constraints amount to unification problems that can always be eagerly solved.
-The linear \picalc{} however requires that some constraints are deferred to a later stage, when more information becomes available elsewhere.
+In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
+Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we use part of $a$ to send a unit, then send $a$ away over $x$ and terminate.
+When the process terminates, all multiplicities must be unrestricted, but we cannot decide whether a multiplicity must be instantiated to $0$ or to $\omega$.
+Similarly, we cannot know \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process does with $a$.
+However, we have to keep track of the fact that $a$ has been used once to send.
 
-Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we send the unit over $a$, then send $a$ over $x$ and terminate.
-By the time we reach $\send$ both $x$ and $a$ must be unrestricted, and thus channel $x$ must be of some type $\tchan~\tzero~\tone~t$.
-Since we send a unit over $a$, $a$ must be of some type $s$ such that $\tSplit{s}{\tchan~\tzero~\tone~\tunit}{t}$.
-From this we can infer that both $s$ and $t$ must be a channel type and thus $\tSplit{s_i}{\tzero}{t_i}$ and $\tSplit{s_o}{\tone}{t_o}$ for some $s_i$, $t_i$, $s_o$ and $t_o$.
-We cannot however infer these usage annotations: all we can deduce is their lower bounds $\tzero$ and $\tone$.
-The resolution of these constraints must be deferred until more information becomes available in the context of this open process.
-
-In \autoref{constraint-collection} we introduce typing constraints, define substitution and constraint satisfaction, and provide a typing inference algorithm.
+In \autoref{constraint-collection} we introduce typing constraints, define kind-preserving substitution and constraint satisfaction, and provide a typing inference algorithm.
 We show that satisfying the constraints generated by typing inference is enough to make a process typable (soundness), and postulate that for any typable process typing inference will find the most general set of constraints that makes the process typable (completeness).
@@
 In \autoref{constraint-resolution} we solve equality constraints by unification and context split constraints by decomposition.
-Constraints have either no solution, a unique solution, or multiple possible solutions.
-(Committing to a non-unique solution can prove to be the wrong choice in an open process, and thus we avoid doing so.
-Closed processes can have non-unique solutions safely instantiated.)
+Constraints have either no solution, a most general solution, or multiple possible solutions.
+(Committing to a solution that is not the most general can prove to be the wrong choice in an open process, and thus we avoid doing so.
+Closed processes can have solutions that are not the most general safely instantiated.)
 Solving a set of constraints results in a set of substitutions and a set of simplified constraints where those substitutions have already been performed.
 We postulate that satisfying the simplified constraints amounts to satisfying the original constraints after substitution (soundness), and that for every substitution that solves the original constraints we can find a most general substitution that will solve the simplified constraints (completeness).
-Additionally, we postulate that the simplified set of constraints only contains unsatisfiable or non-uniquely satisfiable constraints (progress).
+Additionally, we postulate that the simplified set of constraints only contains constraints that are either unsatisfiable or not the most general.
 
@@
 
-A substitution $\tSubst{\gamma}{\delta}$ maps usage annotation and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
+A kind-preserving substitution $\tSubst{\gamma}{\delta}$ maps usage annotation and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
 The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
@@
 
+% Context splits are solved by creating an extra fresh context together with constraints that bind the three.
+% 
 We encode type inference as a function that takes processes with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$.
@@
 Simplified constraints have either no solution, or multiple possible solutions.
-Equality constraints are solved by unification following McBride \cite{McBride03}, and are either unsatisfiable or uniquely satisfied --- and thus resulting in a substitution.
-Context split constraints are solved recursively until a base case is reached.
+Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of McBride's unification by structural recursion \cite{McBride03}, and have either no solution, or a most general solution that results in a substitution.
+Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they either have a most general solution or not.
 $$
@@
 
-To keep us from returning the original constraints as output we postulate that every constraint we return is either unsatisfiable or not uniquely satisfied.
+To keep us from returning the original constraints as output we postulate that every constraint we return is either unsatisfiable or does not have a most general unifier.
 Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$ we have that:
 \begin{flalign*}
-&\forall c \to c \in cs_2 \to \type{IsNotUnique}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
+&\forall c \to c \in cs_2 \to \type{NotMostGeneral}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
 \end{flalign*}
@@
 \begin{flalign*}
-& \type{IsNotUnique}~c && \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 && \\
+& \type{NotMostGeneral}~c && \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 && \\
 & \type{IsUnsat}~c && \triangleq \neg \exists \sigma. ~ \interpr{\subst{\sigma}{c}} &&

### 2021-05-20 paper/paper.tex
Another pass

@@
 \begin{abstract}
-  To type check a \picalc{} with linear and shared types without relying on type annotations, newly created channels must have their types inferred.
-  However, processes can partially use a channel and then send it away: in this case, the receiving process might further constraint its inferred type.
-  This results in non-local constraints that do not have a most general unifier but have to be kept around.
+  Type checking a \picalc{} with linear and shared types without relying on type annotations entails inferring the types of newly created channels.
+  But in the \picalc{} with linear and shared types (also with graded types) processes can partially use a channel and then send it away: in this case, the receiving process might further constraint the channel's inferred type.
+  This means that there are non-local constraints without a most general unifier that we have to keep around.
 
-  Following \cite{Padovani15}, we provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints.
-  We then solve (some of) these constraints using well-known unification algorithms \cite{McBride03} to find the most general unifier, and defer those that do not have one.
-  We state clear soundness and completeness theorems for both these stages and a progress theorem that ensures that only constraints without a most general unifier are deferred.
+  We approach this problem following Padovani \cite{Padovani15}: we provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints, then solve (some of) these constraints using well-known unification algorithms \cite{McBride03} to find the most general unifier, and defer those constraints that do not have one.
+  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only constraints without a most general unifier are deferred.
   The totality of this work is being mechanised in Agda.
@@
 
-The \picalc{} \todo{cite} models concurrent processing by boiling down process interaction to the transmission of data over communication channels --- where channels too are sent as payload.
+The \picalc{} \todo{cite} models concurrent processing by boiling down concurrent interaction to the transmission of data over communication channels --- where channels too are sent as payload.
 The linear \picalc{} \todo{cite} introduces a type system that ensures that every channel is used \emph{exactly} once.
 This restriction ensures communication privacy, communication safety, and the absence of race conditions.
-We generalise and admit a \picalc{} where the input and output multiplicities of a channel are either $0$ (cannot be used), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
+We generalise and admit a \picalc{} where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
 
-To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within the process.
-To do so we can either ask the user for type annotations \todo{cite}, or try to synthesize these types by looking at how the channels are used.
-The latter approach involves dealing with non-local constraints: a channel can be partially used and then sent away --- the process that receives it might further constrain its type.
-This means that only those constraints with a most general solution can safely be eagerly solved, otherwise we risk over-constraining down the line.
-We use co-contextual typing inference \cite{ErdwegBKKM15} to infer a typing context and gather constraints as we traverse processes bottom-up.
-Typing constraints with a most general unifier are eagerly solved and result in substitutions on the typing context, the remaining typing constraints is carried along.
-Collecting constraints bottom-up has the additional advantage of making typing inference easily parallelisable.
-Empowered with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$.
+To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within a process.
+To do so we can either ask the user for type annotations \todo{cite}, or we can synthesize types by looking at how channels are used.
+We follow this latter approach by traversing processes bottom-up while keeping a typing context containing \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing inference \cite{ErdwegBKKM15}.
+Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
+Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: substituting by a solution that is not the most general risks over-constraining the problem down the line.
+Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint) amounts to emitting the extra constraint $\Gamma = \Delta$.
 
-To the best of our knowledge, this problem is only treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.
+To the best of our knowledge, this problem has only been treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.
 However, his work does not appear to be mechanised, and metatheoretical properties like soundness and completeness are only informally addressed.
@@
 On top, we define a standard type system with linear and shared types using context-splits.
-We then provide an overview of how type inference works, show how constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}) are defined, and state that both phases are sound and complete with regards to the previously defined type system.
+We then provide an overview of how typing inference works, show how constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}) are defined, and state that both phases are sound and complete with regards to the previously defined type system.
 
-We have proven that constraint collection is sound, and that equality constraint resolution is sound (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms).
-The remainder proofs are still in progress.
-
-(Note on notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
+We have proven that constraint collection is sound, and adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms, that the resolution of equality constraints is sound.
+The remaining proofs are still in progress.
+(A note on notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
 
@@
 We define a standard syntax and type system for the linear \picalc{}.
-The only non-standard feature is that types allow for \emph{metavariables} (\emph{holes}) within them.
+The only non-standard feature is that types allow for \emph{metavariables} within them.
 
@@
 Variable references $i_n$ are of type $\Fin{n}$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
-\todo{add type assertions}
 \[
@@
 
-Both types and usage annotations contain \emph{metavariables} (\emph{holes}).
-We use a common set of metavariables for both types and usage annotations -- this makes their handling uniform.
+Both types and usage annotations contain metavariables.
+We use a common set of metavariables for both --- this makes their handling uniform.
 A context of kinds \(\gamma\) keeps track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$.
 We refer to a metavariable of kind \(k\) in a kinding context \(\gamma\) as $\tvar{\gamma}{k}$.
-We can now define usage annotations and types in one go (note that we define $\tsum$ and $\tprod$ together for brevity):
+We can now define usage annotations and types in one go (we define $\tzero$, $\tone$ $\tomega$ at once for brevity):
 \begin{mathpar}
@@
   \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
-  {\tsum~s~t ~|~ \tprod~s~t : \tkind{\gamma}{\ktype}}
+  {\tprod~s~t : \tkind{\gamma}{\ktype}}
 
-  \inferrule { } {\tzero : \tkind{\gamma}{\kusage}}
+  \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
+  {\tsum~s~t : \tkind{\gamma}{\ktype}}
 
-  \inferrule { } {\tone : \tkind{\gamma}{\kusage}}
-
-  \inferrule { } {\tomega : \tkind{\gamma}{\kusage}}
+  \inferrule { } {\tzero ~|~ \tone ~|~ \tomega : \tkind{\gamma}{\kusage}}
 \end{mathpar}
-For convenience we henceforth use $\ttype{\gamma}$ to stand for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ to stand for $\tkind{\gamma}{\kusage}$.
+We henceforth use $\ttype{\gamma}$ for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ for $\tkind{\gamma}{\kusage}$.
 
@@
 We define context splits $\tSplit{\Gamma}{\Delta}{\Theta}$ pointwise on types.
-Splits on types are defined recursively pointwise on usage annotations -- note however that only the usage annotations on the top level of a channel (that is, not within its payload) are split.
-Splits on usage annotations are defined as follows -- note that they are not necessarily unique:
+Splits on types are defined recursively pointwise on usage annotations --- note however that only the usage annotations at the top of a channel are split, not those within its payload.
+Splits on usage annotations are defined as follows:
 \begin{mathpar}
@@
 \end{mathpar}
-Following \cite{Padovani15}, we say a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ --- (similarly $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$, respectively).
+Following \cite{Padovani15}, we say that a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ --- (similarly $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$, respectively).
 
@@
 \label{typing-judgments}
-We can finally define our typing judgment for variables, expressions and processes -- some are omitted for brevity.
+We can finally define our typing judgment for variables, expressions and processes (some are omitted for brevity).
 
@@
 
-Co-contextual type inference traverses processes bottom-up collecting constraints.
-This has the advantage of making inference subproblems independent and thus easily parallelisable.
+Co-contextual typing inference traverses processes bottom-up keeping a typing context and collecting constraints.
+This makes inference subproblems independent and thus easily parallelisable.
 In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
 Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we use part of $a$ to send a unit, then send $a$ away over $x$ and terminate.
-When the process terminates, all multiplicities must be unrestricted, but we cannot decide whether a multiplicity must be instantiated to $0$ or to $\omega$.
-Similarly, we cannot know \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process does with $a$.
-However, we have to keep track of the fact that $a$ has been used once to send.
+When the process terminates, all multiplicities must be unrestricted, but we cannot commit to either instantiation $\tzero$ or $\tomega$.
+Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.
+Nonetheless, we do have to keep track of the fact that $a$ has been used once to send.
 
-In \autoref{constraint-collection} we introduce typing constraints, define kind-preserving substitution and constraint satisfaction, and provide a typing inference algorithm.
+In \autoref{constraint-collection} we introduce typing constraints, and define kind-preserving substitution, constraint satisfaction, and a typing inference algorithm.
 We show that satisfying the constraints generated by typing inference is enough to make a process typable (soundness), and postulate that for any typable process typing inference will find the most general set of constraints that makes the process typable (completeness).
@@
 In \autoref{constraint-resolution} we solve equality constraints by unification and context split constraints by decomposition.
-Constraints have either no solution, a most general solution, or multiple possible solutions.
-(Committing to a solution that is not the most general can prove to be the wrong choice in an open process, and thus we avoid doing so.
-Closed processes can have solutions that are not the most general safely instantiated.)
-Solving a set of constraints results in a set of substitutions and a set of simplified constraints where those substitutions have already been performed.
+Constraints with a most general solution can be solved into substitutions.
+(Constraints without a most general solution can be solved by instantiation once the process is closed.)
+Solving a set of constraints thus results in a set of substitutions and a set of simplified constraints where those substitutions have already been performed.
 We postulate that satisfying the simplified constraints amounts to satisfying the original constraints after substitution (soundness), and that for every substitution that solves the original constraints we can find a most general substitution that will solve the simplified constraints (completeness).
-Additionally, we postulate that the simplified set of constraints only contains constraints that are either unsatisfiable or not the most general.
+Additionally, we postulate that the simplified set of constraints only contains constraints without a most general solution.
 
@@
 
-Constraints of type $\tConstr{\gamma}$ are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, both usage annotations and types.
+Constraints of type $\tConstr{\gamma}$ are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, on both usage annotations and types.
 They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
-Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter can \emph{sometimes} not, and must be deferred.
+Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter can \emph{sometimes} not.
 We use $\tConstrs{\gamma}$ to refer to lists of constraints of type $\tConstr{\gamma}$.
@@
 
-A kind-preserving substitution $\tSubst{\gamma}{\delta}$ maps usage annotation and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
+A kind-preserving substitution $\tSubst{\gamma}{\delta}$ maps usage annotations and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
 The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
@@
 
-We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$.
+We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$, respectively.
 
@@
 % 
-We encode type inference as a function that takes processes with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on those metavariables $\gamma$.
-We define a similar function for type inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression:
+We encode typing inference as a function that takes a process with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on metavariables $\gamma$.
+We define a similar function for typing inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression:
 \begin{flalign*}
@@
 \end{flalign*}
-These functions are total: if processes and expressions are untypable, their constraints will simply be unsolvable.
+These functions are total: if processes and expressions are untypable, their constraints will be unsolvable.
 Let us provide a couple of examples illustrating how this function is defined:
 
-For the expression $\sfst~e$ we first recursively infer $e$: $\func{inferExpr}~e = \gamma,\Gamma,Cs,\tau$.
-We then add two fresh metavariables $?t$ and $?s$ of kind $\ktype$ to $\gamma$, add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $cs$, then return $?t$ as the type of $e$.
+For the expression $\sfst~e$ we
+\begin{enumerate*}[label=\arabic*)]
+  \item recursively infer $e$ resulting in $\gamma,\Gamma,Cs,\tau$;
+  \item add two fresh type metavariables $?t$ and $?s$ to $\gamma$;
+  \item add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $Cs$; and
+  \item return $?t$ as the type of $\sfst~e$.
+\end{enumerate*}
 
-For the expression $\scase~e~p~q$ we first recursively infer $p$ and $q$ and $e$: $\func{inferProc}~p = \theta_p, (\Theta_p, s) , Cs_p$, $\func{inferProc}~q = \theta_q, (\Theta_q, t) , Cs_q$, and $\func{inferExpr}~e = \delta_e, \Delta_e, Cs_e, \tau$.
-We create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$.
-We take the union of the metavariables in $\gamma$, $\theta_p$, $\theta_q$ and $\delta_e$
-We take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tsum~s~t}$ and return $\Gamma$ as the inferred context.
+For the expression $\scase~e~p~q$ we
+  \begin{enumerate*}[label=\arabic*)]
+  \item infer $p$ resulting in $\theta_p, (\Theta_p, s) , Cs_p$;
+  \item infer $q$ resulting in $\theta_q, (\Theta_q, t) , Cs_q$;
+  \item infer $e$ resulting in $\delta_e, \Delta_e, Cs_e, \tau$;
+  \item create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$;
+  \item take the union of the metavariables in $\gamma$, $\theta_p$, $\theta_q$ and $\delta_e$;
+  \item take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tsum~s~t}$; and
+  \item return $\Gamma$ as the inferred context.
+\end{enumerate*}
 
@@
 
-Every substitution \(\sigma\) that makes the constraints hold will make the process typable under the substituted context. Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$ we have that:
-\begin{flalign*}
-& (\sigma : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma}{cs}} \to \tProc{(\subst{\sigma}{\Gamma})}{p} &&
-\end{flalign*}
+Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $p$ typable under $\subst{\sigma}{\Gamma}$.
 
@@
 
-For every context \(\Delta\) that makes the process typable there exists a most general substitution \(\sigma\) that will solve the constraints and which substitutes in \(\Gamma\) to something more general than \(\Delta\).
-We define $\Delta ~ \func{\subseteq} ~ \Gamma$ as $\exists \delta. ~ \exists (\sigma : \tSubst{\_}{\delta}). ~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}$.
-Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$ we have that:
-\begin{flalign*}
-  &(\Delta : \tCtx{n}{\gamma}) \to \tProc{\Delta}{p} \to && \\
-  & \exists \delta. ~ \exists (\sigma : \tSubst{\gamma}{\delta}). ~ \interpr{\subst{\sigma}{cs}} \times \Delta \subseteq (\subst{\sigma}{\Gamma}) &&
-\end{flalign*}
+Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $\subst{\sigma}{\Gamma}$ (we define $\Delta ~ \func{\subseteq} ~ \Gamma$ as $\exists \delta.
+~ \exists \sigma .
+~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}$).
 
@@
 Solving a set of constraints results in a set of substitutions and an unsolved set of simplified constraints where those substitutions have already been performed.
-Simplified constraints have either no solution, or multiple possible solutions.
+The constraints that have been left unsolved do not have a most general solution.
 Constraints of the form $\eqconstr{x}{y}$ are solved by unification using a kinded version of McBride's unification by structural recursion \cite{McBride03}, and have either no solution, or a most general solution that results in a substitution.
-Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they either have a most general solution or not.
+Constraints of the form $\sumconstr{x}{y}{z}$ are solved recursively until a base case is reached, at which point they either have a most general solution or they do not.
 $$
@@
 
-Satisfying the simplified constraints satisfies the original constraints once substitutions are applied: $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$ we have that:
-\[
-\begin{aligned}
-& (\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_2}} \to \interpr{\subst{\sigma_f \cdot \sigma}{cs_1}}
-\end{aligned}
-\]
+Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$, every substitution $\sigma_f$ that satisfies the simplified constraints ($\interpr{\subst{\sigma_f}{cs_2}}$) satisfies the original constraints after substitutions are applied ($\interpr{\subst{\sigma_f}{(\subst{\sigma}{cs_1})}}$).
 
@@
 
-Any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold will be a specialisation \(\sigma_g\) of the returned substitution \(\sigma\).
-The specialisation \(\sigma_g\) makes the returned constraints \(cs_2\) hold.
-Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$ we have that:
-\begin{flalign*}
-&(\sigma_f : \tSubst{\gamma}{\delta}) \to \interpr{\subst{\sigma_f}{cs_1}} \to && \\
-&\exists (\sigma_g : \tSubst{\_}{\delta}). ~ \tEq{\sigma_f}{\sigma_g \cdot \sigma} \times \interpr{\subst{\sigma_g}{cs_2}} &&
-\end{flalign*}
+Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$, any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold ($\interpr{\subst{\sigma_f}{cs_1}}$) can be decomposed into $\sigma$ followed by a certain \(\sigma_g\) ($\sigma_f ~ \type{\doteq} ~ \sigma_g \cdot \sigma$) that makes the returned constraints \(cs_2\) hold ($\interpr{\subst{\sigma_g}{cs_2}}$).
 
@@
 
-To keep us from returning the original constraints as output we postulate that every constraint we return is either unsatisfiable or does not have a most general unifier.
-Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$ we have that:
-\begin{flalign*}
-&\forall c \to c \in cs_2 \to \type{NotMostGeneral}~c ~ \type{\uplus} ~ \type{IsUnsat}~c &&
-\end{flalign*}
-where
-\begin{flalign*}
-& \type{NotMostGeneral}~c && \triangleq \exists \sigma_1. \interpr{\subst{\sigma_1}{c}} \times \exists \sigma_2. ~ \interpr{\subst{\sigma_2}{c}} \times \sigma_1 \not\doteq \sigma_2 && \\
-& \type{IsUnsat}~c && \triangleq \neg \exists \sigma. ~ \interpr{\subst{\sigma}{c}} &&
-\end{flalign*}
+Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ we return has a most general unifier, where a most general unifier for a constraint $c$ is defined as $\exists \sigma.
+~ \interpr{\subst{\sigma}{c}} \times (\forall \sigma_f .
+~ \interpr{\subst{\sigma_f}{c}} \times (\exists \sigma_g.
+~ \sigma_f ~ \type{\doteq} ~ \sigma_g \cdot \sigma))$.
 

### 2021-05-20 paper/paper.tex
Add missing bibliography

@@
 \usepackage[inline]{enumitem}
-\usepackage{todonotes}
 
@@
 
-The \picalc{} \todo{cite} models concurrent processing by boiling down concurrent interaction to the transmission of data over communication channels --- where channels too are sent as payload.
-The linear \picalc{} \todo{cite} introduces a type system that ensures that every channel is used \emph{exactly} once.
+The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing by boiling down concurrent interaction to the transmission of data over communication channels --- where channels too are sent as payload.
+The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
 This restriction ensures communication privacy, communication safety, and the absence of race conditions.
@@
 To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within a process.
-To do so we can either ask the user for type annotations \todo{cite}, or we can synthesize types by looking at how channels are used.
+To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can synthesize types by looking at how channels are used.
 We follow this latter approach by traversing processes bottom-up while keeping a typing context containing \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing inference \cite{ErdwegBKKM15}.

### 2021-05-20 paper/paper.tex
Rework abstract

@@
 \begin{abstract}
-  Type checking a \picalc{} with linear and shared types without relying on type annotations entails inferring the types of newly created channels.
-  But in the \picalc{} with linear and shared types (also with graded types) processes can partially use a channel and then send it away: in this case, the receiving process might further constraint the channel's inferred type.
-  This means that there are non-local constraints without a most general unifier that we have to keep around.
+  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within the process.
+  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and always eagerly solve typing constraints into substitutions that are then applied to the typing context.
+  However in our setting, typing constraints do not always come with have a most general unifier, and thus cannot always be eagerly solved.
+  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collecting typing constraints too.
+  We then solve those constraints with a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
 
-  We approach this problem following Padovani \cite{Padovani15}: we provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up collecting constraints, then solve (some of) these constraints using well-known unification algorithms \cite{McBride03} to find the most general unifier, and defer those constraints that do not have one.
-  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only constraints without a most general unifier are deferred.
+  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
   The totality of this work is being mechanised in Agda.

### 2021-05-21 paper/paper.tex
Another pass

@@
 \begin{abstract}
-  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within the process.
-  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and always eagerly solve typing constraints into substitutions that are then applied to the typing context.
-  However in our setting, typing constraints do not always come with have a most general unifier, and thus cannot always be eagerly solved.
-  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collecting typing constraints too.
-  We then solve those constraints with a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
+  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within a process.
+  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and eagerly solve typing constraints into substitutions that are then applied to the typing context.
+  However in our setting, typing constraints do not always come with have a most general solution, and thus cannot always be eagerly solved.
 
+  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
+  We then solve those constraints that have a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
   We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
@@
 Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
-Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: substituting by a solution that is not the most general risks over-constraining the problem down the line.
-Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint) amounts to emitting the extra constraint $\Gamma = \Delta$.
+Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
+Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint).
 

### 2021-05-24 paper/paper.tex
Another pass to the abstract

@@
 \begin{abstract}
-  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within a process.
-  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and eagerly solve typing constraints into substitutions that are then applied to the typing context.
-  However in our setting, typing constraints do not always come with have a most general solution, and thus cannot always be eagerly solved.
+  While a \picalc{} with linear types ensures the privacy and safety of concurrent communication, allowing unrestricted non-linear communication is key to model real-world services.
+  To type check a \picalc{} process with linear and shared types without relying on user-provided type annotations we must infer the types of the channels created within the process.
+  If we limit ourselves to the shared \picalc{}, we can traverse a process bottom-up and eagerly solve typing constraints into substitutions that are then applied to the typing context.
+  However in a setting with both linear and shared types, typing constraints do not always come with have a most general solution, and thus cannot always be eagerly solved.
 
@@
   We then solve those constraints that have a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
-  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
+  We state clear soundness and completeness theorems for both these phases, and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
   The totality of this work is being mechanised in Agda.

### 2021-05-24 paper/paper.tex
Another pass

@@
   While a \picalc{} with linear types ensures the privacy and safety of concurrent communication, allowing unrestricted non-linear communication is key to model real-world services.
-  To type check a \picalc{} process with linear and shared types without relying on user-provided type annotations we must infer the types of the channels created within the process.
-  If we limit ourselves to the shared \picalc{}, we can traverse a process bottom-up and eagerly solve typing constraints into substitutions that are then applied to the typing context.
-  However in a setting with both linear and shared types, typing constraints do not always come with have a most general solution, and thus cannot always be eagerly solved.
+  To implement a decision procedure that type checks a \picalc{} process with linear and shared types without relying on user-provided type annotations we must infer the types of the channels created within the process.
+  If we limit ourselves to the shared \picalc{}, we can traverse the process bottom-up and eagerly solve typing constraints into substitutions and apply them to the typing context.
+  However, in a setting with both linear and shared types, typing constraints do not always come with a most general solution, and thus cannot always be eagerly solved.
 
   We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
-  We then solve those constraints that have a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
-  We state clear soundness and completeness theorems for both these phases, and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
+  We then solve those constraints that have a most general solution (by using well-known unification algorithms \cite{McBride03}) while deferring the rest until more information becomes available.
+  We state clear soundness and completeness theorems separating both these phases, and a progress theorem that ensures that only those constraints without a most general solution are deferred.
   The totality of this work is being mechanised in Agda.
@@
 
-The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing by boiling down concurrent interaction to the transmission of data over communication channels --- where channels too are sent as payload.
+The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing, boiling it down to the transmission of data over communication channels --- where channels too are sent as payload.
 The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
 This restriction ensures communication privacy, communication safety, and the absence of race conditions.
-We generalise and admit a \picalc{} where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
+We generalise to a \picalc{} with linear and shared types, where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
 
-To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within a process.
-To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can synthesize types by looking at how channels are used.
-We follow this latter approach by traversing processes bottom-up while keeping a typing context containing \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing inference \cite{ErdwegBKKM15}.
+To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within the process.
+To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can infer types by looking at how channels are used.
+We follow this latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
 Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
-Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
-Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint).
+Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
+Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$, assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint.
 
@@
 We start defining an untyped but well scoped \picalc{} using type-level de Bruijn indices \cite{deBruijn72} and embed a small expression language that handles composite sum and product types.
-On top, we define a standard type system with linear and shared types using context-splits.
-We then provide an overview of how typing inference works, show how constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}) are defined, and state that both phases are sound and complete with regards to the previously defined type system.
+On top, we define a standard type system with linear and shared types using context-splits \autoref{type-system}.
+We then provide an overview of how typing inference works, define constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}), and state that both phases are sound and complete with regards to the type system defined in \autoref{type-system}.
 
-We have proven that constraint collection is sound, and adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms, that the resolution of equality constraints is sound.
+We have proven that constraint collection is sound and (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms) that the resolution of equality constraints is sound too.
 The remaining proofs are still in progress.
@@
 \section{Type System}
+\label{type-system}
 
@@
 \paragraph{Syntax}
-\label{syntax}
 
@@
 We define context splits $\tSplit{\Gamma}{\Delta}{\Theta}$ pointwise on types.
-Splits on types are defined recursively pointwise on usage annotations --- note however that only the usage annotations at the top of a channel are split, not those within its payload.
+Splits on types are defined recursively on usage annotations --- note that only the usage annotations at the top of a channel are split, not those within its payload.
 Splits on usage annotations are defined as follows:
@@
 \end{mathpar}
-Following \cite{Padovani15}, we say that a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ --- (similarly $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$, respectively).
+Following \cite{Padovani15}, a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$).
 
@@
 \label{typing-judgments}
-We can finally define our typing judgment for variables, expressions and processes (some are omitted for brevity).
+We can finally define our typing judgment for variables, expressions and processes (some cases are omitted for brevity).
 
@@
 In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
-Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we use part of $a$ to send a unit, then send $a$ away over $x$ and terminate.
-When the process terminates, all multiplicities must be unrestricted, but we cannot commit to either instantiation $\tzero$ or $\tomega$.
+Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send $a$ away over $x$ and terminate.
+When the process terminates, all multiplicities must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
 Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.
@@
 (Constraints without a most general solution can be solved by instantiation once the process is closed.)
-Solving a set of constraints thus results in a set of substitutions and a set of simplified constraints where those substitutions have already been performed.
+Solving a set of constraints thus results in a set of substitutions and a set of simplified constraints that had those substitutions applied.
 We postulate that satisfying the simplified constraints amounts to satisfying the original constraints after substitution (soundness), and that for every substitution that solves the original constraints we can find a most general substitution that will solve the simplified constraints (completeness).
-Additionally, we postulate that the simplified set of constraints only contains constraints without a most general solution.
+Additionally, we postulate that the simplified set of constraints only contains constraints without a most general solution (progress).
 
@@
 Constraints of type $\tConstr{\gamma}$ are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, on both usage annotations and types.
-They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ are meant to be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ are meant to be added up to $S$.
+They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ must be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ must add up to $S$.
 Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter can \emph{sometimes} not.
@@
 
-We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their proof counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$, respectively.
+We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their reified counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$, respectively.
 
@@
 
-% Context splits are solved by creating an extra fresh context together with constraints that bind the three.
-% 
-We encode typing inference as a function that takes a process with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on metavariables $\gamma$.
-We define a similar function for typing inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$ for the expression:
+Typing inference takes a process with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on metavariables $\gamma$.
+We define a similar function for typing inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$:
 \begin{flalign*}
@@
 \end{flalign*}
-These functions are total: if processes and expressions are untypable, their constraints will be unsolvable.
+These functions are total: if a process or expression is untypable its constraints will be unsolvable.
+
 Let us provide a couple of examples illustrating how this function is defined:
-
 For the expression $\sfst~e$ we
-\begin{enumerate*}[label=\arabic*)]
+\begin{enumerate*}[label=\textcolor{gray}{\arabic*)}]
   \item recursively infer $e$ resulting in $\gamma,\Gamma,Cs,\tau$;
@@
 For the expression $\scase~e~p~q$ we
-  \begin{enumerate*}[label=\arabic*)]
+  \begin{enumerate*}[label=\textcolor{gray}{\arabic*)}]
   \item infer $p$ resulting in $\theta_p, (\Theta_p, s) , Cs_p$;
@@
 
-Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $p$ typable under $\subst{\sigma}{\Gamma}$.
+Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $\tProc{(\subst{\sigma}{\Gamma})}{p}$ hold.
 
@@
 
-Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $\subst{\sigma}{\Gamma}$ (we define $\Delta ~ \func{\subseteq} ~ \Gamma$ as $\exists \delta.
-~ \exists \sigma .
-~ \tEq{\Delta}{(\subst{\sigma}{\Gamma})}$).
+Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $(\subst{\sigma}{\Gamma})$ --- a specialisation $\Delta ~ \func{\subseteq} ~ (\subst{\sigma}{\Gamma})$ is defined as $\exists \sigma_f .
+~ \tEq{\Delta}{(\subst{\sigma_f}{(\subst{\sigma}{\Gamma})})}$.
 
@@
 
-Solving a set of constraints results in a set of substitutions and an unsolved set of simplified constraints where those substitutions have already been performed.
+Solving a set of constraints results in a set of substitutions and an unsolved set of simplified constraints where those substitutions have already been applied.
 The constraints that have been left unsolved do not have a most general solution.
@@
 
-Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ we return has a most general unifier, where a most general unifier for a constraint $c$ is defined as $\exists \sigma.
+Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ has a most general solution, where a most general solution for a constraint $c$ is defined as $\exists \sigma.
 ~ \interpr{\subst{\sigma}{c}} \times (\forall \sigma_f .

### 2021-05-24 paper/paper.tex
Rework abstract and intro

@@
 \begin{abstract}
-  While a \picalc{} with linear types ensures the privacy and safety of concurrent communication, allowing unrestricted non-linear communication is key to model real-world services.
-  To implement a decision procedure that type checks a \picalc{} process with linear and shared types without relying on user-provided type annotations we must infer the types of the channels created within the process.
-  If we limit ourselves to the shared \picalc{}, we can traverse the process bottom-up and eagerly solve typing constraints into substitutions and apply them to the typing context.
+  A \picalc{} with linear types ensures the privacy and safety of concurrent communication.
+  Allowing unrestricted non-linear communication however is key to model real-world services.
+  To implement a decision procedure that type checks a \picalc{} process with both linear and shared types one can either rely on user-provided type annotations, or infer the types of the channels created within the process.
+  We choose to reduce the burden on the user by following the latter approach.
+  If we limit ourselves to the shared \picalc{}, we can traverse a process bottom-up and eagerly solve typing constraints into substitutions and apply them to the typing context.
   However, in a setting with both linear and shared types, typing constraints do not always come with a most general solution, and thus cannot always be eagerly solved.
 
-  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
+  We provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
   We then solve those constraints that have a most general solution (by using well-known unification algorithms \cite{McBride03}) while deferring the rest until more information becomes available.
@@
 To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can infer types by looking at how channels are used.
-We follow this latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
+To reduce the burden on the user, we follow the latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
 Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.

### 2021-05-24 paper/paper.tex
Add discussion section

@@
 \usepackage{mathpartir}
-\mprset {sep=0.5em} % Horizontal space between premises
+\mprset {sep=0.6em} % Horizontal space between premises
 
@@
 We then provide an overview of how typing inference works, define constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}), and state that both phases are sound and complete with regards to the type system defined in \autoref{type-system}.
-
-We have proven that constraint collection is sound and (adapting McBride's work \cite{McBride03} to deal with intrinsically kinded terms) that the resolution of equality constraints is sound too.
-The remaining proofs are still in progress.
 (A note on notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
@@
 %             { \tVar{\Gamma}{i}{t} }
-
   \inferrule { \Gamma : \tCtx{n}{\gamma} \\ \tun{\Gamma} \\ t : \ttype{\gamma} }
@@
              % \rulename{suc}
-
 %  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ e : \sExpr{n} \\ t : \ttype{\gamma}}
@@
              % \rulename{var}
-
-  \inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} }
+  %\inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} }
              % \rulename{unit}
@@
              % \rulename{inl}
-
-  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{s} \\ \tExpr{\Theta}{f}{t} }
-             { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
+  %\inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{s} \\ \tExpr{\Theta}{f}{t} }
+  %           { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
              %\rulename{pair}
-
 %  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ p : \sProc{n}}
@@
              %\rulename{new}
-
-  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tchan~\tone~\tzero~t} \\ \tProc{\Theta,t}{p} }
-             { \tProc{\Gamma}{\srecv~e~p} }
-             %\rulename{recv}
+  %\inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tchan~\tone~\tzero~t} \\ \tProc{\Theta,t}{p} }
+  %           { \tProc{\Gamma}{\srecv~e~p} }
+  %           %\rulename{recv}
 
@@
 
+\section{Discussion}
+
+We have outlined a procedure for decidable type checking and inference of a \picalc{} with linear and shared types.
+Constraint collection and constraint resolution are kept separate, and their metatheory allows for deferred constraints.
+We have proven in Agda that constraint collection and the resolution of equality constraints is sound, the remaining proofs are still in progress.
+
 \bibliographystyle{abbrvnat}

### 2021-05-24 paper/paper.tex
Be consistent with use of shared/unrestricted/non-linear

@@
   A \picalc{} with linear types ensures the privacy and safety of concurrent communication.
-  Allowing unrestricted non-linear communication however is key to model real-world services.
+  Allowing shared (unlimited) communication however is key to model real-world services.
   To implement a decision procedure that type checks a \picalc{} process with both linear and shared types one can either rely on user-provided type annotations, or infer the types of the channels created within the process.
@@
   We state clear soundness and completeness theorems separating both these phases, and a progress theorem that ensures that only those constraints without a most general solution are deferred.
-  The totality of this work is being mechanised in Agda.
+  This work is being mechanised in Agda.
 \end{abstract}
@@
 This restriction ensures communication privacy, communication safety, and the absence of race conditions.
-We generalise to a \picalc{} with linear and shared types, where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unrestricted use).
+We generalise to a \picalc{} with linear and shared types, where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).
 
@@
 \end{mathpar}
-Following \cite{Padovani15}, a context $\Gamma$ is unrestricted (shared, non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$).
+Following \cite{Padovani15}, a context $\Gamma$ is unrestricted (non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$).
 

### 2021-05-24 paper/paper.tex
Use theorem envs, refer to theorems in discussion

@@
 
+\theoremstyle{definition}\newtheorem{mytheorem}{Theorem}[section]
 \renewcommand*{\sectionautorefname}{\S\!\!\,}
@@
 
-\paragraph{Inference Soundness}
+\begin{mytheorem}[Inference Soundness]\label{inference-soundness}
+Given $\func{infer}~p$ returns $\gamma , \Gamma , cs$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $\tProc{(\subst{\sigma}{\Gamma})}{p}$ hold.
+\end{mytheorem}
 
-Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $\tProc{(\subst{\sigma}{\Gamma})}{p}$ hold.
-
-\paragraph{Inference Completeness}
-
-Given $\tEq{\func{infer}~p}{\gamma , \Gamma , cs}$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $(\subst{\sigma}{\Gamma})$ --- a specialisation $\Delta ~ \func{\subseteq} ~ (\subst{\sigma}{\Gamma})$ is defined as $\exists \sigma_f .
+\begin{mytheorem}[Inference Completeness]\label{inference-completeness}
+Given $\func{infer}~p$ returns $\gamma , \Gamma , cs$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $(\subst{\sigma}{\Gamma})$ --- a specialisation $\Delta ~ \func{\subseteq} ~ (\subst{\sigma}{\Gamma})$ is defined as $\exists \sigma_f .
 ~ \tEq{\Delta}{(\subst{\sigma_f}{(\subst{\sigma}{\Gamma})})}$.
+\end{mytheorem}
 
@@
 
-\paragraph{Resolution Soundness}
+\begin{mytheorem}[Resolution Soundness]\label{resolution-soundness}
+Given $\func{solve}~cs_1$ returns $(\sigma, cs_2)$, every substitution $\sigma_f$ that satisfies the simplified constraints ($\interpr{\subst{\sigma_f}{cs_2}}$) satisfies the original constraints after substitutions are applied ($\interpr{\subst{\sigma_f}{(\subst{\sigma}{cs_1})}}$).
+\end{mytheorem}
 
-Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$, every substitution $\sigma_f$ that satisfies the simplified constraints ($\interpr{\subst{\sigma_f}{cs_2}}$) satisfies the original constraints after substitutions are applied ($\interpr{\subst{\sigma_f}{(\subst{\sigma}{cs_1})}}$).
+\begin{mytheorem}[Resolution Completeness]\label{resolution-completeness}
+Given $\func{solve}~cs_1$ returns $(\sigma, cs_2)$, any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold ($\interpr{\subst{\sigma_f}{cs_1}}$) can be decomposed into $\sigma$ followed by a certain \(\sigma_g\) ($\sigma_f ~ \type{\doteq} ~ \sigma_g \cdot \sigma$) that makes the returned constraints \(cs_2\) hold ($\interpr{\subst{\sigma_g}{cs_2}}$).
+\end{mytheorem}
 
-\paragraph{Resolution Completeness}
-
-Given $\tEq{\func{solve}~cs_1}{(\sigma, cs_2)}$, any substitution \(\sigma_f\) that makes the original constraints \(cs_1\) hold ($\interpr{\subst{\sigma_f}{cs_1}}$) can be decomposed into $\sigma$ followed by a certain \(\sigma_g\) ($\sigma_f ~ \type{\doteq} ~ \sigma_g \cdot \sigma$) that makes the returned constraints \(cs_2\) hold ($\interpr{\subst{\sigma_g}{cs_2}}$).
-
-\paragraph{Resolution Progress}
-
-Given $\func{solve}~cs_1 \equiv (\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ has a most general solution, where a most general solution for a constraint $c$ is defined as $\exists \sigma.
+\begin{mytheorem}[Resolution Progress]\label{resolution-progress}
+Given $\func{solve}~cs_1$ returns $(\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ has a most general solution, where a most general solution for a constraint $c$ is defined as $\exists \sigma.
 ~ \interpr{\subst{\sigma}{c}} \times (\forall \sigma_f .
@@
 ~ \sigma_f ~ \type{\doteq} ~ \sigma_g \cdot \sigma))$.
+\end{mytheorem}
 
@@
 Constraint collection and constraint resolution are kept separate, and their metatheory allows for deferred constraints.
-We have proven in Agda that constraint collection and the resolution of equality constraints is sound, the remaining proofs are still in progress.
+We have proved in Agda the soundness of constraint collection \autoref{inference-soundness} and of equality constraint resolution \autoref{resolution-soundness}, the remaining proofs are still in progress.
 

### 2021-07-17 paper/paper.tex
Apply some of the minor changes suggested in reviews

@@
 \newcommand{\tSubst}[2]{\type{Subst}~#1~#2}
-\newcommand{\interpr}[1]{\func{[\![} #1 \func{]\!]}}
+\newcommand{\interpr}[1]{\func{\llbracket} #1 \func{\rrbracket}}
 
@@
 Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
-Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
+Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
 Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$, assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint.
@@
 A context of kinds \(\gamma\) keeps track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$.
-We refer to a metavariable of kind \(k\) in a kinding context \(\gamma\) as $\tvar{\gamma}{k}$.
-We can now define usage annotations and types in one go (we define $\tzero$, $\tone$ $\tomega$ at once for brevity):
+We define $\tvar{\gamma}{k}$ as the set of variables of kind \(k\) in a kinding context \(\gamma\), and $\tkind{\gamma}{k}$ as the set of types \(\ktype\) and usage annotations \(\kusage\) (we omit unit and sum types and define $\tzero$, $\tone$, $\tomega$ at once for brevity):
 \begin{mathpar}
@@
 
-  \inferrule { }
-  {\tunit : \tkind{\gamma}{\ktype}}
-
   \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
   {\tprod~s~t : \tkind{\gamma}{\ktype}}
-
-  \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
-  {\tsum~s~t : \tkind{\gamma}{\ktype}}
 
@@
 We define context splits $\tSplit{\Gamma}{\Delta}{\Theta}$ pointwise on types.
-Splits on types are defined recursively on usage annotations --- note that only the usage annotations at the top of a channel are split, not those within its payload.
-Splits on usage annotations are defined as follows:
+Splits on types are defined pointwise on usage annotations (but are not applied to a channel's payload) and splits on usage annotations are defined as follows:
 \begin{mathpar}
@@
   \inferrule { } {\tSplit{\tomega}{x}{y}}
+
+  \inferrule {\tSplit{i_x}{i_y}{i_z} \\ \tSplit{o_x}{o_y}{o_z}} {\tSplit{\tchan~i_x~o_x~t}{\tchan~i_y~o_y~t}{\tchan~i_z~o_z~t}}
 \end{mathpar}
@@
 
-A kind-preserving substitution $\tSubst{\gamma}{\delta}$ maps usage annotations and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall k \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$.
-The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall k \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
+A kind-preserving substitution $\tSubst{\gamma}{\delta}$ maps usage annotations and type metavariables in $\gamma$ to usage annotations and types in $\delta$, that is, $\forall \{k\} \to \tvar{\gamma}{k} \to \tkind{\delta}{k}$ for an implicit $k$.
+The function $\subst{\sigma}{t}$ of type $\tSubst{\gamma}{\delta} \to (\forall \{k\} \to \tkind{\gamma}{k} \to \tkind{\delta}{k})$ performs the substitution by replacing all the metavariables in $t$ with their corresponding terms in $\sigma$.
 Substitutions on constraints are defined pointwise on their arguments.
@@
 Typing inference takes a process with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on metavariables $\gamma$.
-We define a similar function for typing inference on expressions, this time also returning a type $\tau$ of type $\ttype{\gamma}$:
+We define a similar function for typing inference on expressions, this time also returning a type in $\ttype{\gamma}$:
 \begin{flalign*}
@@
   \begin{enumerate*}[label=\textcolor{gray}{\arabic*)}]
-  \item infer $p$ resulting in $\theta_p, (\Theta_p, s) , Cs_p$;
-  \item infer $q$ resulting in $\theta_q, (\Theta_q, t) , Cs_q$;
-  \item infer $e$ resulting in $\delta_e, \Delta_e, Cs_e, \tau$;
+  \item infer $p$ resulting in $(\theta_p, \Theta_p, s , Cs_p)$;
+  \item infer $q$ resulting in $(\theta_q, \Theta_q, t , Cs_q)$;
+  \item infer $e$ resulting in $(\delta_e, \Delta_e, Cs_e, \tau)$;
   \item create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$;

### 2021-07-27 paper/paper.tex
Finish with minor review changes

@@
 The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
-This restriction ensures communication privacy, communication safety, and the absence of race conditions.
-We generalise to a \picalc{} with linear and shared types, where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).
+This restriction ensures communication safety, and the absence of race conditions.
+We generalise to a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).
 
@@
 We define a standard syntax for the \picalc{} using type-level de Bruijn indices.
-Variable references $i_n$ are of type $\Fin{n}$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
+Variable references $i_{1+n}$ are of type $\Fin{(1 + n)}$ with constructors $\fzero$ and $\fsuc~i_n$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
 \[
@@
 \label{typing-judgments}
-We can finally define our typing judgment for variables, expressions and processes (some cases are omitted for brevity).
+We can now define our typing judgment for variables ($\tVar{\Gamma}{i}{t}$, where variable $i$ under context $\Gamma$ is of type $t$), expressions ($\tExpr{\Gamma}{e}{t}$, where expression $e$ under context $\Gamma$ is well typed with type $t$) and processes ($\tProc{\Gamma}{p}$, where process $p$ is well typed under context $\Gamma$).
+Some cases are omitted for brevity.
 
@@
 Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send $a$ away over $x$ and terminate.
-When the process terminates, all multiplicities must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
+When the process terminates, all usage annotations must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
 Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.

### 2021-07-27 paper/paper.tex
Use pseudo BNF for kinding judgment

@@
 Both types and usage annotations contain metavariables.
-We use a common set of metavariables for both --- this makes their handling uniform.
+To make their handling uniform we use a common set of metavariables for both.
 A context of kinds \(\gamma\) keeps track of whether a metavariable is of type kind $\ktype$ or usage annotation kind $\kusage$.
-We define $\tvar{\gamma}{k}$ as the set of variables of kind \(k\) in a kinding context \(\gamma\), and $\tkind{\gamma}{k}$ as the set of types \(\ktype\) and usage annotations \(\kusage\) (we omit unit and sum types and define $\tzero$, $\tone$, $\tomega$ at once for brevity):
-\begin{mathpar}
-  \inferrule {m : \tvar{\gamma}{k}} {\tmvar~m : \tkind{\gamma}{k}}
-
-  \inferrule {
-    i : \tkind{\gamma}{\kusage} \\
-    o : \tkind{\gamma}{\kusage} \\
-    t : \tkind{\gamma}{\ktype}}
-  {\tchan~i~o~t : \tkind{\gamma}{\ktype}}
-
-  \inferrule {s : \tkind{\gamma}{\ktype} \\ t : \tkind{\gamma}{\ktype} }
-  {\tprod~s~t : \tkind{\gamma}{\ktype}}
-
-  \inferrule { } {\tzero ~|~ \tone ~|~ \tomega : \tkind{\gamma}{\kusage}}
-\end{mathpar}
-We henceforth use $\ttype{\gamma}$ for $\tkind{\gamma}{\ktype}$ and $\tusage{\gamma}$ for $\tkind{\gamma}{\kusage}$.
+Variable references $m$ are of type $\tvar{\gamma}{k}$: the set of variables of kind \(k\) in a kinding context \(\gamma\).
+Usage annotations $i$ and $o$ are of type $\tkind{\gamma}{\kusage}$, and types $s$ and $t$ of type $\tkind{\gamma}{\ktype}$.
+We henceforth abbreviate $\tkind{\gamma}{\ktype}$ as $\ttype{\gamma}$ and $\tkind{\gamma}{\kusage}$ as $\tusage{\gamma}$.
+\[
+\begin{aligned}[c]
+  i ~ o :=& ~ \tmvar_{\kusage} ~ m ~|~ \tzero ~|~ \tone ~|~ \tomega \\
+  s ~ t :=& ~ \tmvar_{\ktype} ~ m ~|~ \tunit ~|~ \tchan ~ i ~ o ~ t ~|~ \tprod ~ s ~ t ~|~ \tsum ~ s ~ t
+\end{aligned}
+\]
 

### 2021-07-27 paper/paper.tex
Add more interesting typing rules
Rewrite most of the constraint collection section

@@
 \usepackage[inline]{enumitem}
+\usepackage{todonotes}
 
@@
 \newcommand{\rulename}[1]{{\tiny \textsc{(#1)}}}
+\newcommand{\var}[1]{\textcolor{black}{\mathnormal{#1}}}
 \newcommand{\constr}[1]{\textcolor{olive}{\mathtt{#1}}}
@@
 \newcommand{\tmvar}{\constr{mvar}}
-\newcommand{\tchan}{\constr{chan}}
+\newcommand{\tchan}[3]{\constr{chan}_{\constr{[}#1\constr{,}#2\constr{]}} ~ #3}
 \newcommand{\tunit}{\constr{unit}}
@@
 \newcommand{\sumconstr}[3]{\constr{[} ~ #1 ~ \constr{\stackrel{c}{=}} ~ #2 ~ \constr{+} ~ #3 ~ \constr{]}}
+\newcommand{\varconstr}[3]{\constr{[} ~ #1 ~ \constr{\stackrel{c}{\ni}_{\var{#2}}} ~ #3 ~ \constr{]}}
 
@@
 To reduce the burden on the user, we follow the latter approach: we traverse processes bottom-up while keeping a typing context with \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing \cite{ErdwegBKKM15}.
-Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
+Keeping a strictly bottom-up information flow has the additional advantage of making typing inference parallelisable.
 Constraints with a most general solution are solved into substitutions and applied to the typing context, constraints without must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
@@
 \begin{itemize}
-  \item State and prove clear soundness and completeness theorems for both constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}).
+  \item State and prove clear soundness and completeness theorems for both constraint collection and constraint resolution (\autoref{inference}).
   \item Mechanise in Agda the totality of this work.
@@
 We start defining an untyped but well scoped \picalc{} using type-level de Bruijn indices \cite{deBruijn72} and embed a small expression language that handles composite sum and product types.
-On top, we define a standard type system with linear and shared types using context-splits \autoref{type-system}.
-We then provide an overview of how typing inference works, define constraint collection (\autoref{constraint-collection}) and constraint resolution (\autoref{constraint-resolution}), and state that both phases are sound and complete with regards to the type system defined in \autoref{type-system}.
-(A note on notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, \textcolor{gray}{functions are gray}, and we are sorry this is not a poem.)
+On top, we define a standard type system with linear and shared types using context-splits (\autoref{type-system}).
+We then provide an overview of how typing inference works, define constraint collection and constraint resolution (\autoref{inference}), and state that both phases are sound and complete with regards to the type system defined in \autoref{type-system}.
+(Notation: variables are black, \textcolor{blue}{types are blue}, \textcolor{olive}{constructors are green}, and \textcolor{gray}{functions are gray}.)
 
@@
   i ~ o :=& ~ \tmvar_{\kusage} ~ m ~|~ \tzero ~|~ \tone ~|~ \tomega \\
-  s ~ t :=& ~ \tmvar_{\ktype} ~ m ~|~ \tunit ~|~ \tchan ~ i ~ o ~ t ~|~ \tprod ~ s ~ t ~|~ \tsum ~ s ~ t
+  s ~ t :=& ~ \tmvar_{\ktype} ~ m ~|~ \tunit ~|~ \tchan{i}{o}{t} ~|~ \tprod ~ s ~ t ~|~ \tsum ~ s ~ t
 \end{aligned}
@@
 
-  \inferrule {\tSplit{i_x}{i_y}{i_z} \\ \tSplit{o_x}{o_y}{o_z}} {\tSplit{\tchan~i_x~o_x~t}{\tchan~i_y~o_y~t}{\tchan~i_z~o_z~t}}
+  \inferrule {\tSplit{i_x}{i_y}{i_z} \\ \tSplit{o_x}{o_y}{o_z}} {\tSplit{\tchan{i_x}{o_x}{t}}{\tchan{i_y}{o_y}{t}}{\tchan{i_z}{o_z}{t}}}
 \end{mathpar}
@@
 We can now define our typing judgment for variables ($\tVar{\Gamma}{i}{t}$, where variable $i$ under context $\Gamma$ is of type $t$), expressions ($\tExpr{\Gamma}{e}{t}$, where expression $e$ under context $\Gamma$ is well typed with type $t$) and processes ($\tProc{\Gamma}{p}$, where process $p$ is well typed under context $\Gamma$).
-Some cases are omitted for brevity.
+Note that, although fully linear tensor products ($\tprod ~ s ~ t$) cannot directly be eliminated through their projections ($\sfst$ rule), context splits enable their usage annotations to be distributed before projection.
+(Some cases are omitted for brevity.)
 
 \begin{mathpar}
-%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ i : \Fin{n} \\ t : \ttype{\gamma} }
-%             { \tVar{\Gamma}{i}{t} }
   \inferrule { \Gamma : \tCtx{n}{\gamma} \\ \tun{\Gamma} \\ t : \ttype{\gamma} }
              { \tVar{\Gamma,t}{\fzero}{t} }
-             % \rulename{zero}
 
@@
              { \tVar{\Gamma,s}{\fsuc~i}{t} }
-             % \rulename{suc}
-%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ e : \sExpr{n} \\ t : \ttype{\gamma}}
-%             { \tExpr{\Gamma}{e}{t} }
-
-  \inferrule { \tVar{\Gamma}{i}{t} } { \tExpr{\Gamma}{\svar~i}{t} }
-             % \rulename{var}
-  %\inferrule { \tun{\Gamma} } { \tExpr{\Gamma}{\sunit}{\tunit} }
-             % \rulename{unit}
 
   \inferrule { \tExpr{\Gamma}{e}{\tprod~t~s} \\ \tun{s}} {\tExpr{\Gamma}{\sfst~e}{t}}
-             % \rulename{fst}
 
-  \inferrule { \tExpr{\Gamma}{e}{t} } { \tExpr{\Gamma}{\sinl~e}{\tsum~t~s}}
-             % \rulename{inl}
-  %\inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{s} \\ \tExpr{\Theta}{f}{t} }
-  %           { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
-             %\rulename{pair}
-%  \inferrule { \Gamma : \tCtx{n}{\gamma} \\ p : \sProc{n}}
-%             { \tProc{\Gamma}{p} }
+  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{s} \\ \tExpr{\Theta}{f}{t} }
+             { \tExpr{\Gamma}{\spair~e~f}{\tprod~s~t}}
 
   \inferrule { \tun{\Gamma} } { \tProc{\Gamma}{\send} }
-             %\rulename{end}
-
-  \inferrule { \tProc{\Gamma}{p} \\ \tun{\Gamma} }
-             { \tProc{\Gamma}{\srec~p} }
-             %\rulename{rec}
 
   \inferrule { t : \ttype{\gamma} \\ \tProc{\Gamma , t}{p} } { \tProc{\Gamma}{\snew~p} }
-             %\rulename{new}
-  %\inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tchan~\tone~\tzero~t} \\ \tProc{\Theta,t}{p} }
-  %           { \tProc{\Gamma}{\srecv~e~p} }
-  %           %\rulename{recv}
 
-  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tsum~s~t} \\ \tProc{\Theta,s}{p} \\ \tProc{\Theta,t}{q} }
-             { \tProc{\Gamma}{\scase~e~p~q} }
-             %\rulename{case}
+  \inferrule { \tSplit{\Gamma}{\Delta}{\Theta} \\ \tExpr{\Delta}{e}{\tchan{\tone}{\tzero}{t}} \\ \tProc{\Theta,t}{p} }
+             { \tProc{\Gamma}{\srecv~e~p} }
+
+  \mprset{flushleft}
+  \inferrule {
+    \tSplit{\Gamma}{\Delta}{\Theta} \\\\
+    \tSplit{\Theta}{\Psi}{\Xi} \\
+    \tExpr{\Delta}{f}{\tchan{\tzero}{\tone}{t}} \\
+    \tExpr{\Psi}{e}{t} \\
+    \tProc{\Xi}{p}
+  }
+  { \tProc{\Gamma}{\ssend~f~e~p} }
 \end{mathpar}
@@
 
-Co-contextual typing inference traverses processes bottom-up keeping a typing context and collecting constraints.
-This makes inference subproblems independent and thus easily parallelisable.
-In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
-Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send $a$ away over $x$ and terminate.
-When the process terminates, all usage annotations must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
-Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.
-Nonetheless, we do have to keep track of the fact that $a$ has been used once to send.
-
-In \autoref{constraint-collection} we introduce typing constraints, and define kind-preserving substitution, constraint satisfaction, and a typing inference algorithm.
-We show that satisfying the constraints generated by typing inference is enough to make a process typable (soundness), and postulate that for any typable process typing inference will find the most general set of constraints that makes the process typable (completeness).
-
-In \autoref{constraint-resolution} we solve equality constraints by unification and context split constraints by decomposition.
-Constraints with a most general solution can be solved into substitutions.
-(Constraints without a most general solution can be solved by instantiation once the process is closed.)
-Solving a set of constraints thus results in a set of substitutions and a set of simplified constraints that had those substitutions applied.
-We postulate that satisfying the simplified constraints amounts to satisfying the original constraints after substitution (soundness), and that for every substitution that solves the original constraints we can find a most general substitution that will solve the simplified constraints (completeness).
-Additionally, we postulate that the simplified set of constraints only contains constraints without a most general solution (progress).
-
-\subsection{Constraint Collection}
-\label{constraint-collection}
-
 \paragraph{Constraints}
@@
 Constraints of type $\tConstr{\gamma}$ are defined on arguments of type $\tkind{\gamma}{k}$ for some $k$ --- that is, on both usage annotations and types.
-They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ must be unified, and the ternary $\sumconstr{S}{T}{R}$, where $T$ and $R$ must add up to $S$.
-Constraints of the former form can \emph{always} be eagerly solved, while constraints of the latter can \emph{sometimes} not.
+They take two forms: the binary $\eqconstr{S}{T}$, where $S$ and $T$ must be unified, and the ternary $\sumconstr{S}{T}{R}$, where $S$ is split into $T$ and $R$.
+Constraints are pointwise lifted to typing contexts, and we abbreviate with $\varconstr{\Gamma}{i}{t}$ a constraint that states that $i$ must be of type $t$ in $\Gamma$, and all other variables in $\Gamma$ must be unrestricted.
 We use $\tConstrs{\gamma}$ to refer to lists of constraints of type $\tConstr{\gamma}$.
@@
 
-We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into their reified counterparts $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$, respectively.
+We use a $\interpr{\_}$ function to interpret constraints $\eqconstr{S}{T}$ and $\sumconstr{S}{T}{R}$ into claims of their satisfiability $\tEq{S}{T}$ and $\tSplit{S}{T}{R}$, respectively.
 
@@
 Typing inference takes a process with $n$ free variables and returns a metavariable context $\gamma$, a typing context with $n$ free variables containing metavariables in $\gamma$, and a list of constraints on metavariables $\gamma$.
-We define a similar function for typing inference on expressions, this time also returning a type in $\ttype{\gamma}$:
+We define a similar function for typing inference on expressions, this time also returning a type in $\ttype{\gamma}$.
+These functions are total: if a process or expression is untypable its constraints are unsolvable.
 \begin{flalign*}
@@
 \end{flalign*}
-These functions are total: if a process or expression is untypable its constraints will be unsolvable.
 
-Let us provide a couple of examples illustrating how this function is defined:
-For the expression $\sfst~e$ we
-\begin{enumerate*}[label=\textcolor{gray}{\arabic*)}]
-  \item recursively infer $e$ resulting in $\gamma,\Gamma,Cs,\tau$;
-  \item add two fresh type metavariables $?t$ and $?s$ to $\gamma$;
-  \item add constraints $\eqconstr{\tau}{\tprod~?t~?s}$ and $\sumconstr{?s}{?s}{?s}$ to $Cs$; and
-  \item return $?t$ as the type of $\sfst~e$.
-\end{enumerate*}
+The reason we return unsolved constraints is that unlike in the shared \picalc{} where constraints have always a most general unifier, in the shared \emph{and} linear \picalc{} metavariables can be under-constrained.
+Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send whatever is left of $a$ away over $x$ and terminate.
+Let us step through a working example of how inference runs:
+\begin{enumerate}
+  \item on $\send$, inference creates a fresh typing context $\Gamma_0$ with metavariables for $x$ and $a$, and constraints demanding that these should be unrestricted.
+  \item on $\ssend ~ x ~ a$, inference creates fresh typing contexts $\Gamma_4$, $\Gamma_3$, $\Gamma_2$ and $\Gamma_1$, a fresh metavariable $?t$, and constraints $\sumconstr{\Gamma_4}{\Gamma_3}{\Gamma_2}$, $\varconstr{\Gamma_3}{x}{\tchan{\tzero}{\tone}{?t}}$, $\sumconstr{\Gamma_2}{\Gamma_1}{\Gamma_0}$, and $\varconstr{\Gamma_1}{a}{?t}$.
+  \item on $\ssend ~ a ~ \sunit$, inference creates fresh typing contexts $\Gamma_6$, and $\Gamma_5$, and constraints $\sumconstr{\Gamma_6}{\Gamma_5}{\Gamma_4}$, and $\varconstr{\Gamma_5}{a}{\tchan{\tzero}{\tone}{\tunit}}$.
+\end{enumerate}
 
-For the expression $\scase~e~p~q$ we
-  \begin{enumerate*}[label=\textcolor{gray}{\arabic*)}]
-  \item infer $p$ resulting in $(\theta_p, \Theta_p, s , Cs_p)$;
-  \item infer $q$ resulting in $(\theta_q, \Theta_q, t , Cs_q)$;
-  \item infer $e$ resulting in $(\delta_e, \Delta_e, Cs_e, \tau)$;
-  \item create a new metavariable context $\gamma$ and a new typing context $\Gamma$ where every type is a fresh metavariable in $\gamma$;
-  \item take the union of the metavariables in $\gamma$, $\theta_p$, $\theta_q$ and $\delta_e$;
-  \item take the union of the constraints in $Cs_p$, $Cs_q$ and $Cs_e$, and create new constraints $\eqconstr{\Theta_p}{\Theta_q}$, $\sumconstr{\Gamma}{\Delta_e}{\Theta_p}$, and $\eqconstr{\tau}{\tsum~s~t}$; and
-  \item return $\Gamma$ as the inferred context.
-\end{enumerate*}
+Two reasons prevent us from finding most general solutions:
+On one hand, \emph{unrestrictedness} constraints are satisfied by either usage annotations $\tzero$ and $\tomega$.
+This can be overcome by a subtyping relation on usage annotations.
+Additionally, \emph{usage polymorphism} on $?t$ makes inference under-constrained.
+The overall constraint on the metavariable for $a$ is that it can be split into $\tchan{\tzero}{\tone}{\tunit}$, into $?t$, and into some unrestricted leftovers.
+While $?t$ can certainly be eagerly substituted by a channel type $\tchan{?i}{?o}{\tunit}$ for some $?i$ and $?o$, we cannot determine what $?i$ and $?o$ must be.
+The problem here is that we must keep track of the partial usage of $a$ while at the same time allowing the type of $a$ to be polymorphic in the type of $x$.
+As a result, we must keep these under-constrained constraints around (potentially until the process is closed and they can be solved by instantiation) and state meta-theoretical properties abstracting over constraint satisfaction.
 
 \begin{mytheorem}[Inference Soundness]\label{inference-soundness}
-Given $\func{infer}~p$ returns $\gamma , \Gamma , cs$, every substitution \(\sigma\) that makes the constraints $cs$ hold makes $\tProc{(\subst{\sigma}{\Gamma})}{p}$ hold.
+Given $\func{infer}~p$ returns $(\gamma , \Gamma , cs)$, every substitution \(\sigma\) satisfying $\interpr{\subst{\sigma}{cs}}$ makes $\tProc{(\subst{\sigma}{\Gamma})}{p}$ hold.
 \end{mytheorem}
@@
 \begin{mytheorem}[Inference Completeness]\label{inference-completeness}
-Given $\func{infer}~p$ returns $\gamma , \Gamma , cs$, for every context \(\Delta\) that makes $p$ typable, there exists a substitution \(\sigma\) that solves the constraints $cs$ such that $\Delta$ is a specialisation of $(\subst{\sigma}{\Gamma})$ --- a specialisation $\Delta ~ \func{\subseteq} ~ (\subst{\sigma}{\Gamma})$ is defined as $\exists \sigma_f .
+Given $\func{infer}~p$ returns $(\gamma , \Gamma , cs)$, for every context \(\Delta\) such that $\tProc{\Delta}{p}$, there exists a substitution \(\sigma\) satisfying $\interpr{\subst{\sigma}{cs}}$ such that $\Delta$ is a specialisation of $(\subst{\sigma}{\Gamma})$ --- a specialisation $\Delta ~ \func{\subseteq} ~ (\subst{\sigma}{\Gamma})$ is defined as $\exists \sigma_f .
 ~ \tEq{\Delta}{(\subst{\sigma_f}{(\subst{\sigma}{\Gamma})})}$.

### 2021-07-27 paper/paper.tex
Another pass at inference

@@
 
-The reason we return unsolved constraints is that unlike in the shared \picalc{} where constraints have always a most general unifier, in the shared \emph{and} linear \picalc{} metavariables can be under-constrained.
+Unlike in the shared \picalc{}, where constraints have always a most general unifier, in the shared \emph{and} linear \picalc{} metavariables can be under-constrained.
 Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send whatever is left of $a$ away over $x$ and terminate.
@@
   \item on $\send$, inference creates a fresh typing context $\Gamma_0$ with metavariables for $x$ and $a$, and constraints demanding that these should be unrestricted.
-  \item on $\ssend ~ x ~ a$, inference creates fresh typing contexts $\Gamma_4$, $\Gamma_3$, $\Gamma_2$ and $\Gamma_1$, a fresh metavariable $?t$, and constraints $\sumconstr{\Gamma_4}{\Gamma_3}{\Gamma_2}$, $\varconstr{\Gamma_3}{x}{\tchan{\tzero}{\tone}{?t}}$, $\sumconstr{\Gamma_2}{\Gamma_1}{\Gamma_0}$, and $\varconstr{\Gamma_1}{a}{?t}$.
-  \item on $\ssend ~ a ~ \sunit$, inference creates fresh typing contexts $\Gamma_6$, and $\Gamma_5$, and constraints $\sumconstr{\Gamma_6}{\Gamma_5}{\Gamma_4}$, and $\varconstr{\Gamma_5}{a}{\tchan{\tzero}{\tone}{\tunit}}$.
+  \item on $\ssend ~ x ~ a$, inference creates fresh typing contexts $\Gamma_4$, $\Gamma_3$, $\Gamma_2$ and $\Gamma_1$, a fresh metavariable $?t$, and constraints $\sumconstr{\Gamma_4}{\Gamma_3}{\Gamma_2}$, $\varconstr{\Gamma_3}{x}{\tchan{\tzero}{\tone}{?t}}$, $\sumconstr{\Gamma_2}{\Gamma_1}{\Gamma_0}$, and $\varconstr{\Gamma_1}{a}{?t}$, following the typing rules.
+  \item on $\ssend ~ a ~ \sunit$, inference creates fresh typing contexts $\Gamma_6$ and $\Gamma_5$, and constraints $\sumconstr{\Gamma_6}{\Gamma_5}{\Gamma_4}$, and $\varconstr{\Gamma_5}{a}{\tchan{\tzero}{\tone}{\tunit}}$.
 \end{enumerate}
 
-Two reasons prevent us from finding most general solutions:
-On one hand, \emph{unrestrictedness} constraints are satisfied by either usage annotations $\tzero$ and $\tomega$.
-This can be overcome by a subtyping relation on usage annotations.
-Additionally, \emph{usage polymorphism} on $?t$ makes inference under-constrained.
-The overall constraint on the metavariable for $a$ is that it can be split into $\tchan{\tzero}{\tone}{\tunit}$, into $?t$, and into some unrestricted leftovers.
-While $?t$ can certainly be eagerly substituted by a channel type $\tchan{?i}{?o}{\tunit}$ for some $?i$ and $?o$, we cannot determine what $?i$ and $?o$ must be.
-The problem here is that we must keep track of the partial usage of $a$ while at the same time allowing the type of $a$ to be polymorphic in the type of $x$.
-As a result, we must keep these under-constrained constraints around (potentially until the process is closed and they can be solved by instantiation) and state meta-theoretical properties abstracting over constraint satisfaction.
+Here \emph{usage polymorphism} on $?t$ makes inference under-constrained and prevents us from finding a most general solution:
+the constraints on $a$'s type demand that it must be split into $\tchan{\tzero}{\tone}{\tunit}$, into $?t$, and into some unrestricted leftovers, and while $?t$ can eagerly be substituted by a channel type $\tchan{?i}{?o}{\tunit}$ for some $?i$ and $?o$, it is polymorphic in its usage annotations $?i$ and $?o$.
+In other words, we must keep track of the partial usage of $a$ while allowing $x$ to be polymorphic in the type of $a$.
+As a result, these partial usage constraints must be kept around (potentially until the process is closed and they can be solved by instantiation) and meta-theoretical properties must therefore be abstracted over constraint satisfaction.
 

### 2021-07-27 paper/paper.tex
Comments on usage annotations

@@
 The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
-This restriction ensures communication safety, and the absence of race conditions.
-We generalise to a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).
+This restriction ensures communication safety and the absence of race conditions.
+We generalise to a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either usage $0$ (cannot be used to transmit), usage $1$ (must be used exactly once), or usage $\omega$ (unlimited use).
+\footnote{
+  Why support a usage $0$ instead of removing the variable from context?
+it allows the syntax to be independent from the type system, and it allows for a \emph{polarised} treatment of channels, where the same channel variable is used for both input and output.
+  }
 
@@
 \end{mathpar}
+Note that e.g., both $\tSplit{\tzero}{\tzero}{\tzero}$ and $\tSplit{\tomega}{\tzero}{\tzero}$ are possible: we use context splits to allow the type system to lose granularity.
 Following \cite{Padovani15}, a context $\Gamma$ is unrestricted (non-linear) $\tun{\Gamma}$ if it can be split into itself: $\tSplit{\Gamma}{\Gamma}{\Gamma}$ (respectively $\tun{x}$ and $\tun{t}$ for usage annotations $x$ and types $t$).
@@
 We can now define our typing judgment for variables ($\tVar{\Gamma}{i}{t}$, where variable $i$ under context $\Gamma$ is of type $t$), expressions ($\tExpr{\Gamma}{e}{t}$, where expression $e$ under context $\Gamma$ is well typed with type $t$) and processes ($\tProc{\Gamma}{p}$, where process $p$ is well typed under context $\Gamma$).
-Note that, although fully linear tensor products ($\tprod ~ s ~ t$) cannot directly be eliminated through their projections ($\sfst$ rule), context splits enable their usage annotations to be distributed before projection.
-(Some cases are omitted for brevity.)
+Note that, although fully linear tensor products ($\tprod ~ s ~ t$) cannot directly be eliminated through their projections ($\sfst$ rule), context splits enable their usage annotations to be distributed beforehand.
+Similarly, while it may appear that $\srecv$ and $\ssend$ only work on channels with usages ($\tone$, $\tzero$) and vice versa, the context splits in these rules allow usages to lose granularity (and thus accept e.g., ($\tomega$, $\tomega$)).
 

### 2021-07-28 paper/paper.tex
Fix remarks regarding Kobayashi and Padovani's work

@@
 The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing, boiling it down to the transmission of data over communication channels --- where channels too are sent as payload.
-The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
-This restriction ensures communication safety and the absence of race conditions.
-We generalise to a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either usage $0$ (cannot be used to transmit), usage $1$ (must be used exactly once), or usage $\omega$ (unlimited use).
+Type systems for the \picalc{} that support linearity \cite{KobayashiPT96} ensure that linear channels are used \emph{exactly once}, which guarantees communication safety and the absence of race conditions.
+We follow this line of work with a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either usage $0$ (cannot be used to transmit), usage $1$ (must be used exactly once), or usage $\omega$ (unlimited use).
 \footnote{
-  Why support a usage $0$ instead of removing the variable from context?
-it allows the syntax to be independent from the type system, and it allows for a \emph{polarised} treatment of channels, where the same channel variable is used for both input and output.
+  Why support a usage $0$ instead of removing the variable from context altogether?
+it allows the syntax to be independent from the type system, and moreover for a \emph{polarised} treatment of channels, where the same channel variable is used for both input and output.
   }
@@
 
-To the best of our knowledge, this problem has only been treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.
-However, his work does not appear to be mechanised, and metatheoretical properties like soundness and completeness are only informally addressed.
+To the best of our knowledge, this problem has not yet been mechanised, and has only been treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.
 With this work, we aim to:

### 2021-07-28 paper/paper.tex
Add comments on related work

@@
 
+Decidable (and certified) type checking procedures have been provided for a range of linear type systems \cite{Allais15,Allais17,OrchardLE19,Brady21}.
+These type systems however do not deal with types containing usage annotations nested within: in our case, a channel with \emph{payload} type ``channel with input usage 0'' differs in type from a channel with payload type ``channel with input usage 1''.
+As a result, we must roll typing inference and linearity inference into one, and deal with the type polymorphism that arises as a result of usage polymorphism.
 To the best of our knowledge, this problem has not yet been mechanised, and has only been treated in Padovani's work on type reconstruction for composite types \cite{Padovani15}.

### 2021-07-28 paper/paper.tex
Fix typo

@@
 \begin{mytheorem}[Resolution Progress]\label{resolution-progress}
-Given $\func{solve}~cs_1$ returns $(\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ has a most general solution, where a most general solution for a constraint $c$ is defined as $\exists \sigma.
+Given $\func{solve}~cs_1$ returns $(\sigma, cs_2)$, to keep us from returning the original constraints as output (which is both sound and complete), we postulate that none of the constraints $c \in cs_2$ have a most general solution, where a most general solution for a constraint $c$ is defined as $\exists \sigma.
 ~ \interpr{\subst{\sigma}{c}} \times (\forall \sigma_f .
